---
title: "project_1_code"
format: html
editor: visual
---

# PART 1: DATA PROCESSING

## PARSING VARIABLE VALUES

```{r}
#Reading in packages to current session
library(httr)
library(jsonlite)
library(dplyr)
library(tidyverse)
library(lubridate)
```

*Writing functions to parse variable data.*

```{r}
#Writing functions to get urls and parse them. Note that the parsed_data objects on their own are very large and take a long time to load. It's easier to use str() to look at them.
list_vars_21 <- function(api_url1){
  api_url1 <- 'https://api.census.gov/data/2021/acs/acs1/pums/variables.json'
  data_url1 <- httr::GET(api_url1)
  parsed_data1 <- fromJSON(rawToChar(data_url1$content))
}
variables_2021 <- list_vars_21(api_url1)

list_vars_22 <- function(api_url2){
  api_url2 <- 'https://api.census.gov/data/2022/acs/acs1/pums/variables.json'
  data_url2 <- httr::GET(api_url2)
  parsed_data2 <- fromJSON(rawToChar(data_url2$content))
}
variables_2022 <- list_vars_22(api_url2)

list_vars_23 <- function(api_url3){
  api_url3 <- 'https://api.census.gov/data/2023/acs/acs1/pums/variables.json'
  data_url3 <- httr::GET(api_url3)
  parsed_data3 <- fromJSON(rawToChar(data_url3$content))
}
variables_2023 <- list_vars_23(api_url3)

list_vars_24 <- function(api_url4){
  api_url4 <- 'https://api.census.gov/data/2024/acs/acs1/pums/variables.json'
  data_url4 <- httr::GET(api_url4)
  parsed_data4 <- fromJSON(rawToChar(data_url4$content))
}
variables_2024 <- list_vars_24(api_url4)
```

*Putting variable lists into .rds files.*

```{r}
#Previously attempted this by putting everything into one data frame, but it ended up recycling parts due to uneven lengths.
saveRDS(variables_2021, file = '2021_variables.rds')
saveRDS(variables_2022, file = '2022_variables.rds')
saveRDS(variables_2023, file = '2023_variables.rds')
saveRDS(variables_2024, file = '2024_variables.rds')
```

*Testing the functions to see if they work. The functions work with both the link to the variables and the api_url object.*

```{r}
#Must include variables column when retrieving a specific element, otherwise it will return a NULL.
test_21 <- list_vars_21(api_url1)
test_21$variables$ST

test_22 <- list_vars_22(api_url2)
test_22$variables$AGEP

test_23 <- list_vars_23(api_url3)
test_23$variables$GRPIP

test_24 <- list_vars_24(api_url4)
test_24$variables$SEX
```

## PARSING API DATA

*Getting URL, writing helper function, and parsing data. These take a long time to run. (functions work, no editing needed, remove note later)*

```{r}
#Getting 2021 survey
pums_data_21 <- httr::GET('api.census.gov/data/2021/acs/acs1/pums?get=AGEP,GRPIP,JWAP,JWDP,JWMNP,PWGTP,FER,HHL,SCH,SCHL,SEX,ST,REGION,DIVISION&key=416ac733e884381d5fc80215dc879af0edb1dd8b')

#Using function to make it into a tibble
helper_21 <- function(pums_data_21){
  parsed_data <- fromJSON(rawToChar(pums_data_21$content))
  tibble_data <- as_tibble(parsed_data)
}

#Creating named object from tibble and checking structure
parsed_tibble_21 <- helper_21(pums_data_21)
str(parsed_tibble_21)
```

```{r}
#Getting 2022 survey
pums_data_22 <- httr::GET('api.census.gov/data/2022/acs/acs1/pums?get=AGEP,GRPIP,JWAP,JWDP,JWMNP,PWGTP,FER,HHL,SCH,SCHL,SEX,ST,REGION,DIVISION&key=416ac733e884381d5fc80215dc879af0edb1dd8b')

#Using function to make it into a tibble
helper_22 <- function(pums_data_22){
  parsed_data2 <- fromJSON(rawToChar(pums_data_22$content))
  tibble_data2 <- as_tibble(parsed_data2)
}

#Creating named object from tibble and checking structure
parsed_tibble_22 <- helper_22(pums_data_22)
str(parsed_tibble_22)
```

```{r}
#Getting 2023 survey
pums_data_23 <- httr::GET('api.census.gov/data/2023/acs/acs1/pums?get=AGEP,GRPIP,JWAP,JWDP,JWMNP,PWGTP,FER,HHL,SCH,SCHL,SEX,STATE,REGION,DIVISION&key=416ac733e884381d5fc80215dc879af0edb1dd8b')

#Using function to make it into a tibble
helper_23 <- function(pums_data_23){
  parsed_data3 <- fromJSON(rawToChar(pums_data_23$content))
  tibble_data3 <- as_tibble(parsed_data3)
}

#Creating named object from tibble and checking structure
parsed_tibble_23 <- helper_23(pums_data_23)
str(parsed_tibble_23)
```

```{r}
#Getting 2024 survey
pums_data_24 <- httr::GET('api.census.gov/data/2024/acs/acs1/pums?get=AGEP,GRPIP,JWAP,JWDP,JWMNP,PWGTP,FER,HHL,SCH,SCHL,SEX,STATE,REGION,DIVISION&key=416ac733e884381d5fc80215dc879af0edb1dd8b')

#Using function to make it into a tibble
helper_24 <- function(pums_data_24){
  parsed_data4 <- fromJSON(rawToChar(pums_data_24$content))
  tibble_data4 <- as_tibble(parsed_data4)
}

#Creating named object from tibble and checking structure
parsed_tibble_24 <- helper_24(pums_data_24)
str(parsed_tibble_24)
```

*Writing function that allows user to change year of survey. This function takes a long time to run and returns a lot of data. The example at the bottom uses str() to reduce the amount of information returned. (function works, no editing needed, remove note later)*

```{r}
#Writing function to return year of survey
return_date <- function(year_vector){
  year_vector <- c(2021, 2022, 2023, 2024)
  if(2021){
    api_2021 <- 'api.census.gov/data/2021/acs/acs1/pums?get=AGEP,GRPIP,JWAP,JWDP,JWMNP,PWGTP,FER,HHL,SCH,SCHL,SEX,ST,REGION,DIVISION&key=416ac733e884381d5fc80215dc879af0edb1dd8b'
    url_2021 <- httr::GET(api_2021)
    parsed_data_2021 <- fromJSON(rawToChar(url_2021$content))
    date_tibble_21 <- as_tibble(parsed_data_2021)
    return(date_tibble_21)
  }else if(2022){
    api_2022 <- 'api.census.gov/data/2022/acs/acs1/pums?get=AGEP,GRPIP,JWAP,JWDP,JWMNP,PWGTP,FER,HHL,SCH,SCHL,SEX,ST,REGION,DIVISION&key=416ac733e884381d5fc80215dc879af0edb1dd8b'
    url_2022 <- httr::GET(api_2022)
    parsed_data_2022 <- fromJSON(rawToChar(url_2022$content))
    date_tibble_22 <- as_tibble(parsed_data_2022)
    return(date_tibble_22)
  }else if(2023){
    api_2023 <- 'api.census.gov/data/2023/acs/acs1/pums?get=AGEP,GRPIP,JWAP,JWDP,JWMNP,PWGTP,FER,HHL,SCH,SCHL,SEX,STATE,REGION,DIVISION&key=416ac733e884381d5fc80215dc879af0edb1dd8b'
    url_2023 <- httr::GET(api_2023)
    parsed_data_2023 <- fromJSON(rawToChar(url_2023$content))
    date_tibble_23 <- as_tibble(parsed_data_2023)
    return(date_tibble_23)
  }else if(2024){
    api_2024 <- 'api.census.gov/data/2024/acs/acs1/pums?get=AGEP,GRPIP,JWAP,JWDP,JWMNP,PWGTP,FER,HHL,SCH,SCHL,SEX,STATE,REGION,DIVISION&key=416ac733e884381d5fc80215dc879af0edb1dd8b'
    url_2024 <- httr::GET(api_2024)
    parsed_data_2024 <- fromJSON(rawToChar(url_2024$content))
    date_tibble_24 <- as_tibble(parsed_data_2024)
    return(date_tibble_24)
  }
}

test_date <- return_date(2021)
str(test_date)
```

*Writing functions to specify numeric variables to be returned*. (function still needs to be worked on)

```{r}
return_time_variable <- function(year_of_survey){
  year_of_survey <- list(2021, 2022, 2023, 2024)
  if(2021){
    readRDS('2021_variables.rds')
    time_JWAP <- as.numeric(variables_2021$variables$JWAP)
    JWAP_clean <- time_JWAP[!is.na(time_JWAP)]
    time_JWDP <- as.numeric(variables_2021$variables$JWDP)
    JWDP_clean <- time_JWDP[!is.na(time_JWDP)]
    JWAP_midpoint <- lapply(X = JWAP_clean, FUN = mean.difftime, na.rm = TRUE)
    JWDP_midpoint <- lapply(X = JWDP_clean, FUN = mean.difftime, na.rm = TRUE)
    JWAP_time <- hms(JWAP_midpoint)
    JWDP_time <- hms(JWDP_midpoint)
    return(list(JWAP = JWAP_time, JWDP = JWDP_time))
  }else if(2022){
    readRDS('2022_variables.rds')
    time_JWAP <- as.numeric(variables_2022$variables$JWAP)
    JWAP_clean <- time_JWAP[!is.na(time_JWAP)]
    time_JWDP <- as.numeric(variables_2022$variables$JWDP)
    JWDP_clean <- time_JWDP[!is.na(time_JWDP)]
    JWAP_midpoint <- lapply(X = JWAP_clean, FUN = mean.difftime, na.rm = TRUE)
    JWDP_midpoint <- lapply(X = JWDP_clean, FUN = mean.difftime, na.rm = TRUE)
    JWAP_time <- hms(JWAP_midpoint)
    JWDP_time <- hms(JWDP_midpoint)
    return(list(JWAP = JWAP_time, JWDP = JWDP_time))
  }else if(2023){
    readRDS('2023_variables.rds')
    time_JWAP <- as.numeric(variables_2023$variables$JWAP)
    JWAP_clean <- time_JWAP[!is.na(time_JWAP)]
    time_JWDP <- as.numeric(variables_2023$variables$JWDP)
    JWDP_clean <- time_JWDP[!is.na(time_JWDP)]
    JWAP_midpoint <- lapply(X = JWAP_clean, FUN = mean.difftime, na.rm = TRUE)
    JWDP_midpoint <- lapply(X = JWDP_clean, FUN = mean.difftime, na.rm = TRUE)
    JWAP_time <- hms(JWAP_midpoint)
    JWDP_time <- hms(JWDP_midpoint)
    return(list(JWAP = JWAP_time, JWDP = JWDP_time))
  }else if(2024){
    readRDS('2024_variables.rds')
    time_JWAP <- as.numeric(variables_2024$variables$JWAP)
    JWAP_clean <- time_JWAP[!is.na(time_JWAP)]
    time_JWDP <- as.numeric(variables_2024$variables$JWDP)
    JWDP_clean <- time_JWDP[!is.na(time_JWDP)]
    JWAP_midpoint <- lapply(X = JWAP_clean, FUN = mean.difftime, na.rm = TRUE)
    JWDP_midpoint <- lapply(X = JWDP_clean, FUN = mean.difftime, na.rm = TRUE)
    JWAP_time <- hms(JWAP_midpoint)
    JWDP_time <- hms(JWDP_midpoint)
    return(list(JWAP = JWAP_midpoint, JWDP = JWDP_midpoint))
  }else{
    'ERROR'
  }
}

test_time_24 <- return_time_variable(2024)
test_time_24
```

```{r}
test_num <- httr::GET('api.census.gov/data/2021/acs/acs1/pums?get=AGEP,GASP,GRPIP,JWMNP&key=416ac733e884381d5fc80215dc879af0edb1dd8b')
parsed_test <- fromJSON(rawToChar(test_num$content))
tibble_test <- as_tibble(parsed_test)
```

```{r}
return_num_variable <- function(year_of_survey){
  year_of_survey <- c(2021, 2022, 2023, 2024)
  if(2021){
    num_api <- 'api.census.gov/data/2021/acs/acs1/pums?get=AGEP,GASP,GRPIP,JWMNP&key=416ac733e884381d5fc80215dc879af0edb1dd8b'
    get_url <- httr::GET(num_api)
    parsed_url <- fromJSON(rawToChar(get_url$content))
    num_tibble <- as_tibble(parsed_url)
  }else if(2022){
    num_api <- 'api.census.gov/data/2022/acs/acs1/pums?get=AGEP,GASP,GRPIP,JWMNP&key=416ac733e884381d5fc80215dc879af0edb1dd8b'
    get_url <- httr::GET(num_api)
    parsed_url <- fromJSON(rawToChar(get_url$content))
    num_tibble <- as_tibble(parsed_url)
  }else if(2023){
    num_api <- 'api.census.gov/data/2023/acs/acs1/pums?get=AGEP,GASP,GRPIP,JWMNP&key=416ac733e884381d5fc80215dc879af0edb1dd8b'
    get_url <- httr::GET(num_api)
    parsed_url <- fromJSON(rawToChar(get_url$content))
    num_tibble <- as_tibble(parsed_url)
  }else if(2024){
    num_api <- 'api.census.gov/data/2024/acs/acs1/pums?get=AGEP,GASP,GRPIP,JWMNP&key=416ac733e884381d5fc80215dc879af0edb1dd8b'
    get_url <- httr::GET(num_api)
    parsed_url <- fromJSON(rawToChar(get_url$content))
    num_tibble <- as_tibble(parsed_url)
  }
 
}

```

*Writing functions to specify categorical variables to be returned. (functions still need to be worked on)*

```{r}
return_cat_vars <- function(year_of_survey){
  year_of_survey <- c(2021,2022,2023,2024)
  if(2021){
    readRDS('2021_variables.rds')
    FER <- mutate(map(variables_2021$variables$FER, as.factor))
    HHL <- mutate(map(variables_2021$variables$HHL, as.factor))
    SCH <- mutate(map(variables_2021$variables$SCH, as.factor))
    SCHL <- mutate(map(variables_2021$variables$SCHL, as.factor))
    SEX <- mutate(map(variables_2021$variables$SEX, as.factor))
    return(list(FER = FER, HHL = HHL, SCH = SCH, SCHL = SCHL, SEX = SEX))
  }else if(2022){
    readRDS('2022_variables.rds')
    FER <- as_factor(variables_2022$variables$FER)
    HHL <- as_factor(variables_2022$variables$HHL)
    SCH <- as_factor(variables_2022$variables$SCH)
    SCHL <- as_factor(variables_2022$variables$SCHL)
    SEX <- as_factor(variables_2022$variables$SEX)
    return(list(FER = FER, HHL = HHL, SCH = SCH, SCHL = SCHL, SEX = SEX))
  }else if(2023){
    readRDS('2023_variables.rds')
    FER <- as_factor(variables_2023$variables$FER)
    HHL <- as_factor(variables_2023$variables$HHL)
    SCH <- as_factor(variables_2023$variables$SCH)
    SCHL <- as_factor(variables_2023$variables$SCHL)
    SEX <- as_factor(variables_2023$variables$SEX)
    return(list(FER = FER, HHL = HHL, SCH = SCH, SCHL = SCHL, SEX = SEX))
  }else if(2024){
    readRDS('2024_variables.rds')
    FER <- as_factor(variables_2024$variables$FER)
    HHL <- as_factor(variables_2024$variables$HHL)
    SCH <- as_factor(variables_2024$variables$SCH)
    SCHL <- as_factor(variables_2024$variables$SCHL)
    SEX <- as_factor(variables_2024$variables$SEX)
    return(list(FER = FER, HHL = HHL, SCH = SCH, SCHL = SCHL, SEX = SEX))
  }else{
    'ERROR'
  }
}

test_cat_vars <- return_cat_vars(2021)
```

*Writing function to specify geography level. (function works, no editing needed, remove note later)*

```{r}
geography_level <- function(year_geo){
 year_geo <- c(2021, 2022, 2023, 2024)
 if(2021){
   geo_api <- 'api.census.gov/data/2021/acs/acs1/pums?get=ST,REGION,DIVISION&key=416ac733e884381d5fc80215dc879af0edb1dd8b'
   geo_url <- httr::GET(geo_api)
   geo_parsed <- fromJSON(rawToChar(geo_url$content))
   geo_tibble <- as_tibble(geo_parsed)
   return(geo_tibble)
 }else if(2022){
   geo_api <- 'api.census.gov/data/2022/acs/acs1/pums?get=ST,REGION,DIVISION&key=416ac733e884381d5fc80215dc879af0edb1dd8b'
   geo_url <- httr::GET(geo_api)
   geo_parsed <- fromJSON(rawToChar(geo_url$content))
   geo_tibble <- as_tibble(geo_parsed)
   return(geo_tibble)
 }else if(2023){
   geo_api <- 'api.census.gov/data/2023/acs/acs1/pums?get=STATE,REGION,DIVISION&key=416ac733e884381d5fc80215dc879af0edb1dd8b'
   geo_url <- httr::GET(geo_api)
   geo_parsed <- fromJSON(rawToChar(geo_url$content))
   geo_tibble <- as_tibble(geo_parsed)
   return(geo_tibble)
 }else if(2024){
   geo_api <- 'api.census.gov/data/2024/acs/acs1/pums?get=STATE,REGION,DIVISION&key=416ac733e884381d5fc80215dc879af0edb1dd8b'
   geo_url <- httr::GET(geo_api)
   geo_parsed <- fromJSON(rawToChar(geo_url$content))
   geo_tibble <- as_tibble(geo_parsed)
   return(geo_tibble)
 }else{
   'ERROR'
 }
}

test_geo <- geography_level(2021)
test_geo
```

*Writing function to specify region, division, or state to subset data on.*

```{r}
#Default values: Region = South, Division = Puerto Rico, State = 'North Carolina/NC'
```

## COMBINING YEARS

*Writing function to specify multiple years of survey data.*

```{r}

```

# PART TWO: WRITING A GENERIC FUNCTION FOR SUMMARIZING

## NUMERIC SUMMARIES

```{r}
summary.census <- function(tibble, numeric_vector, weight_vector){
  tibble <- 'put tibble here'
  numeric_vector <- 'put numeric vector here'
  weight_vector <- 'put weight vector here'
  sample_mean <- sum(numeric_vector*weight_vector)/sum(weight_vector)
  sample_std_dev <- sqrt(sum(numeric_vectorˆ2*weight_vector)/sum(weight_vector)-sample_meanˆ2)
  return(list(sample_mean, sample_std_dev))
}
```

## GRAPHICAL SUMMARIES

```{r}

```
