---
title: "project_1_code"
format: html
editor: visual
---

# PART 1: DATA PROCESSING

## PARSING VARIABLE VALUES

```{r}
#Reading in packages to current session
library(httr)
library(jsonlite)
library(dplyr)
library(tidyverse)
library(lubridate)
library(stringr)
```

*Writing functions to parse variable data.*

```{r}
#Writing functions to get urls and parse them. Note that the parsed_data objects on their own are very large and take a long time to load. It's easier to use str() to look at them.
list_vars_21 <- function(api_url1){
  api_url1 <- 'https://api.census.gov/data/2021/acs/acs1/pums/variables.json'
  data_url1 <- httr::GET(api_url1)
  parsed_data1 <- fromJSON(rawToChar(data_url1$content))
}
variables_2021 <- list_vars_21(api_url1)

list_vars_22 <- function(api_url2){
  api_url2 <- 'https://api.census.gov/data/2022/acs/acs1/pums/variables.json'
  data_url2 <- httr::GET(api_url2)
  parsed_data2 <- fromJSON(rawToChar(data_url2$content))
}
variables_2022 <- list_vars_22(api_url2)

list_vars_23 <- function(api_url3){
  api_url3 <- 'https://api.census.gov/data/2023/acs/acs1/pums/variables.json'
  data_url3 <- httr::GET(api_url3)
  parsed_data3 <- fromJSON(rawToChar(data_url3$content))
}
variables_2023 <- list_vars_23(api_url3)

list_vars_24 <- function(api_url4){
  api_url4 <- 'https://api.census.gov/data/2024/acs/acs1/pums/variables.json'
  data_url4 <- httr::GET(api_url4)
  parsed_data4 <- fromJSON(rawToChar(data_url4$content))
}
variables_2024 <- list_vars_24(api_url4)
```

*Putting variable lists into .rds files.*

```{r}
#Previously attempted this by putting everything into one data frame, but it ended up recycling parts due to uneven lengths.
saveRDS(variables_2021, file = '2021_variables.rds')
saveRDS(variables_2022, file = '2022_variables.rds')
saveRDS(variables_2023, file = '2023_variables.rds')
saveRDS(variables_2024, file = '2024_variables.rds')
```

*Testing the functions to see if they work. The functions work with both the link to the variables and the api_url object.*

```{r}
#Must include variables column when retrieving a specific element, otherwise it will return a NULL.
test_21 <- list_vars_21(api_url1)
test_21$variables$ST

test_22 <- list_vars_22(api_url2)
test_22$variables$AGEP

test_23 <- list_vars_23(api_url3)
test_23$variables$GRPIP

test_24 <- list_vars_24(api_url4)
test_24$variables$SEX
```

## PARSING API DATA

*Getting URL, writing helper function, and parsing data.*

*These take a long time to run, querying the API directly through a link takes a long time for my computer to process.*

*These presented my first problems with this project, I was getting errors about using both \$ and \[\] within my function to call certain objects. In the as_tibble function, I had been calling the object parsed_data\$data. I had seen in the reading and remembered from previous homework that that's something you should do when reading in API data. I found out that this was the issue when I accidentally wrote a function similar to this one 'wrong' and it ended up working perfectly fine. The part that was different was the lack of '\$data', as I had forgotten it.*

```{r}
#Getting 2021 survey
pums_data_21 <- httr::GET('api.census.gov/data/2021/acs/acs1/pums?get=AGEP,GRPIP,JWAP,JWDP,JWMNP,PWGTP,FER,HHL,SCH,SCHL,SEX,ST,REGION,DIVISION&key=416ac733e884381d5fc80215dc879af0edb1dd8b')

#Using function to make it into a tibble
helper_21 <- function(pums_data_21){
  parsed_data <- fromJSON(rawToChar(pums_data_21$content))
  tibble_data <- as_tibble(parsed_data)
}

#Creating named object from tibble and checking structure
parsed_tibble_21 <- helper_21(pums_data_21)
str(parsed_tibble_21)
```

```{r}
#Getting 2022 survey
pums_data_22 <- httr::GET('api.census.gov/data/2022/acs/acs1/pums?get=AGEP,GRPIP,JWAP,JWDP,JWMNP,PWGTP,FER,HHL,SCH,SCHL,SEX,ST,REGION,DIVISION&key=416ac733e884381d5fc80215dc879af0edb1dd8b')

#Using function to make it into a tibble
helper_22 <- function(pums_data_22){
  parsed_data2 <- fromJSON(rawToChar(pums_data_22$content))
  tibble_data2 <- as_tibble(parsed_data2)
}

#Creating named object from tibble and checking structure
parsed_tibble_22 <- helper_22(pums_data_22)
str(parsed_tibble_22)
```

```{r}
#Getting 2023 survey
pums_data_23 <- httr::GET('api.census.gov/data/2023/acs/acs1/pums?get=AGEP,GRPIP,JWAP,JWDP,JWMNP,PWGTP,FER,HHL,SCH,SCHL,SEX,STATE,REGION,DIVISION&key=416ac733e884381d5fc80215dc879af0edb1dd8b')

#Using function to make it into a tibble
helper_23 <- function(pums_data_23){
  parsed_data3 <- fromJSON(rawToChar(pums_data_23$content))
  tibble_data3 <- as_tibble(parsed_data3)
}

#Creating named object from tibble and checking structure
parsed_tibble_23 <- helper_23(pums_data_23)
str(parsed_tibble_23)
```

```{r}
#Getting 2024 survey
pums_data_24 <- httr::GET('api.census.gov/data/2024/acs/acs1/pums?get=AGEP,GRPIP,JWAP,JWDP,JWMNP,PWGTP,FER,HHL,SCH,SCHL,SEX,STATE,REGION,DIVISION&key=416ac733e884381d5fc80215dc879af0edb1dd8b')

#Using function to make it into a tibble
helper_24 <- function(pums_data_24){
  parsed_data4 <- fromJSON(rawToChar(pums_data_24$content))
  tibble_data4 <- as_tibble(parsed_data4)
}

#Creating named object from tibble and checking structure
parsed_tibble_24 <- helper_24(pums_data_24)
str(parsed_tibble_24)
```

*Writing function that allows user to change year of survey.*

*I previously was using if/else logic to write these functions, in the style below:*

my_function \<- function(year){

year \<- c(2021, 2022, 2023, 2024)

if(2021){

Code to be run

}else if(2022){

Code to be run

} and so on.

*I was having a lot of problems getting the right things to be returned with this. The functions themselves were also extremely long. I tried paste0 after seeing it in the reading for this class and it worked, so I was able to make the functions much shorter. I tried it with just the paste function but it ended up not working due to the space that it added.*

```{r}
#Writing function to return year of survey
return_date <- function(year = 2024){
  rds <- paste0(year, '_variables.rds')
  variables <- readRDS(rds)
  return(variables)
}

#Testing this function
test_date <- return_date(2021)
str(test_date)
```

*Writing functions to change numeric variables from characters to numeric and then return them.*

*With this function, I had the most amount of trouble with the type of objects the variables were. I was continuously getting errors and warnings about all sorts of things, particularly when trying to convert them to numeric variables. Finally, making them into vectors worked.*

```{r}
#Writing function that returns numeric variables, also changes them from characters to numeric
return_num_vars <- function(year, target_var = 'AGEP'){
  rds <- paste0(year, "_variables.rds")
  var_list <- readRDS(rds)
  PWGTP = c(var_list$variables$PWGTP$values$range$min, var_list$variables$PWGTP$values$range$max)
  AGEP = c(var_list$variables$AGEP$values$range$min, var_list$variables$AGEP$values$range$max)
  GASP = c(var_list$variables$GASP$values$range$min, var_list$variables$GASP$values$range$max)
  GRPIP = c(var_list$variables$GRPIP$values$range$min, var_list$variables$GRPIP$values$range$max)
  JWMNP = c(var_list$variables$JWMNP$values$range$min, var_list$variables$JWMNP$values$range$max)
  full_list <- list(AGEP, GASP, GRPIP, JWMNP)
  num_list <- lapply(full_list, as.numeric, na.rm = TRUE)
  return(list(num_list['PWGTP'], num_list[target_var]))
}

test_num <- return_num_vars(year = 2021, target_var = 'GASP')
test_num
```

*Writing a function to turn the time data into numeric time data.*

*This was definitely the part that gave me the most trouble. There was a point where the function I was trying to write would cause RStudio to shut down and become completely unresponsive.*

```{r}
return_JWAP <- function(year){
variables <- paste0(year, '_variables.rds')
variables_list <- readRDS(variables)
variables_df <- data.frame(
  data = unlist(variables_list$variables$JWAP$values$item),
  stringsAsFactors = FALSE
)

to_minutes <- function(time_str, am_pm){
  parts <- str_split(time_str, ':', simplify = TRUE)
  hours <- as.numeric(parts[, 1])
  minutes <- as.numeric(parts[, 2])
  
  is_pm <- str_detect(am_pm, 'p')
  
  hours <- case_when(
    is_pm & hours != 12 ~ hours + 12,
    !is_pm & hours == 12 ~ 0,
    TRUE                 ~ hours
  )
  return(hours * 60 + minutes)
}
processed <- variables_df |>
  mutate(
    start_time = str_extract(data, "^\\d{1,2}:\\d{2}"),
    start_ampm = str_extract(data, "^\\d{1,2}:\\d{2}\\s*[ap]\\.m\\."),
    end_time = str_extract(data, "\\d{1,2}:\\d{2}\\s*[ap]\\.m\\.$") %>% str_extract("^\\d{1,2}:\\d{2}"),
    end_ampm = str_extract(data, "\\d{1,2}:\\d{2}\\s*[ap]\\.m\\.$"),
    start_mins = to_minutes(start_time, start_ampm),
    end_mins = to_minutes(end_time, end_ampm),
    end_mins = if_else(end_mins < start_mins, end_mins + 1440, end_mins),
    midpoint_mins = (start_mins + end_mins) / 2,
    midpoint_mins = midpoint_mins %% 1440,
    midpoint_time = sprintf("%02d:%02d", floor(midpoint_mins / 60), round(midpoint_mins %% 60))
  ) |>
  select(midpoint_time)
print(processed)
}

return_JWDP <- function(year){
  variables <- paste0(year, '_variables.rds')
variables_list <- readRDS(variables)
variables_df <- data.frame(
  data = unlist(variables_list$variables$JWDP$values$item),
  stringsAsFactors = FALSE
)

to_minutes <- function(time_str, am_pm){
  parts <- str_split(time_str, ':', simplify = TRUE)
  hours <- as.numeric(parts[, 1])
  minutes <- as.numeric(parts[, 2])
  
  is_pm <- str_detect(am_pm, 'p')
  
  hours <- case_when(
    is_pm & hours != 12 ~ hours + 12,
    !is_pm & hours == 12 ~ 0,
    TRUE                 ~ hours
  )
  return(hours * 60 + minutes)
}
processed <- variables_df |>
  mutate(
    start_time = str_extract(data, "^\\d{1,2}:\\d{2}"),
    start_ampm = str_extract(data, "^\\d{1,2}:\\d{2}\\s*[ap]\\.m\\."),
    end_time = str_extract(data, "\\d{1,2}:\\d{2}\\s*[ap]\\.m\\.$") %>% str_extract("^\\d{1,2}:\\d{2}"),
    end_ampm = str_extract(data, "\\d{1,2}:\\d{2}\\s*[ap]\\.m\\.$"),
    start_mins = to_minutes(start_time, start_ampm),
    end_mins = to_minutes(end_time, end_ampm),
    end_mins = if_else(end_mins < start_mins, end_mins + 1440, end_mins),
    midpoint_mins = (start_mins + end_mins) / 2,
    midpoint_mins = midpoint_mins %% 1440,
    midpoint_time = sprintf("%02d:%02d", floor(midpoint_mins / 60), round(midpoint_mins %% 60))
  ) |>
  select(midpoint_time)
print(processed)
}

JWAP_test <- return_JWAP(2021)
JWAP_test
JWDP_test <- return_JWDP(2021)
JWDP_test
```

```{r}
#Testing out a way to change the data to the right form
library(hms)
library(lubridate)
test_time <- readRDS('2021_variables.rds')
test_col <- test_time$variables$JWAP$values$item[1]

clean_time <- gsub('to|\\a.m.|\\p.m.|\\:|\\ ', '', test_col)

split_1 <- str_sub(clean_time, start = 1, end = 3)
split_1
split_2 <- str_sub(clean_time, start = 4, end = 6)
split_2
num_1 <- as.numeric(split_1)
num_1
num_2 <- as.numeric(split_2)
num_2
num_vec <- c(num_1, num_2)
mean <- mean(num_vec)
mean
time <- as.POSIXct(sprintf('%04d', mean), format = '%H%M')
time
new_time <- format(time, format = '%H:%M')
new_time
#This does work! This will be used as a reference going forward with writing this function

```

*Writing functions to specify categorical variables to be returned.*

*Writing these didn't give me too much trouble outside of figuring out how to change the variables into factors. I had been previously been using as.factor(), which had not been working.*

```{r}
#Writing function
return_cat_vars <- function(year, target_var = 'SEX'){
rds <- paste0(year, '_variables.rds')
var_list <- readRDS(rds)
FER = factor(unlist(var_list$variables$FER))
HHL = factor(unlist(var_list$variables$HHL))
SCH = factor(unlist(var_list$variables$SCH))
SCHL = factor(unlist(var_list$variables$SCHL))
SEX = factor(unlist(var_list$variables$SEX))
var_list <- list(FER, HHL, SCH, SCHL, SEX)
return(var_list[target_var])
}

#Testing it out
test_cat_vars <- return_cat_vars(2023, 'FER')
test_cat_vars
```

*Writing function to specify geography level.*

*I didn't have much trouble with this one.*

```{r}
#Writing function
geography_level <- function(year, target_var = 'REGION'){
rds <- paste0(year, '_variables.rds')
vars_list <- readRDS(rds)
REGION = vars_list$variables$REGION
STATE = vars_list$variables$STATE
DIVISION = vars_list$variables$DIVISION
geo_list <- list(REGION, STATE, DIVISION)
return(geo_list[target_var])
}

#Testing this function
test_geo <- geography_level(2021)
test_geo
```

*Writing functions to specify region, division, or state to subset data on.*

*This group was very easy to put together.*

```{r}
#Writing the region funcition
subset_region <- function(year){
rds <- paste0(year, '_variables.rds')
vars_list <- readRDS(rds)
region = vars_list$variables$REGION
return(region)
}
#Testing
test_geo <- subset_region(2022)
test_geo
```

```{r}
#Writing the state function

#This one should be used for the 2021 and 2022 data
subset_st <- function(year){
rds <- paste0(year, '_variables.rds')
vars_list <- readRDS(rds)
state = vars_list$variables$ST
return(state)
}

#This one should be used for the 2023 and 2024 data
subset_state <- function(year){
rds <- paste0(year, '_variables.rds')
vars_list <- readRDS(rds)
state = vars_list$variables$STATE
return(state)
}
#Testing
test_st <- subset_st(2021)
test_st

test_state <- subset_state(2024)
test_state
```

```{r}
#Writing the division function
subset_division <- function(year){
rds <- paste0(year, '_variables.rds')
vars_list <- readRDS(rds)
division = vars_list$variables$DIVISION
return(division)
}
#Testing
test_division <- subset_division(2022)
test_division
```

## PUTTING EVERYTHING TOGETHER

*There are two different main functions because of the presence of different variables for state depending on the year.*

```{r}
main_function_21_22 <- function(year){
  rds <- paste0(year, '_variables.rds')
  vars_list <- readRDS(rds)
  final_date <- return_date(year)
  final_num <- return_num_vars(year)
  final_JWAP <- return_JWAP(year)
  final_JWDP <- return_JWDP(year)
  final_cat <- return_cat_vars(year)
  final_geo <- return_geography(year)
  final_region <- subset_region(year)
  final_st <- subset_st(year)
  final_division <- subset_division(year)
  final_list <- list(final_date, final_num, final_JWAP, final_JWDP, final_cat, final_geo, final_region, final_st, final_division, final_list)
  final_tibble <- as_tibble(final_list)
  final_census <- class(final_tibble) <- c("census", class(final_tibble))
  return(final_census)
}

main_function_23_24 <- function(year){
  rds <- paste0(year, '_variables.rds')
  vars_list <- readRDS(rds)
  final_date <- return_date(year)
  final_num <- return_num_vars(year)
  final_JWAP <- return_JWAP(year)
  final_JWDP <- return_JWDP(year)
  final_cat <- return_cat_vars(year)
  final_geo <- return_geography(year)
  final_region <- subset_region(year)
  final_state <- subset_state(year)
  final_division <- subset_division(year)
  final_list <- list(final_date, final_num, final_JWAP, final_JWDP, final_cat, final_geo, final_region, final_state, final_division, final_list)
  final_tibble <- as_tibble(final_list)
  final_census <- class(final_tibble) <- c("census", class(final_tibble))
  return(final_census)
}
```

## COMBINING YEARS

*Writing function to specify multiple years of survey data. This function results in a HUGE amount of returned data. The only issue I had with this is that I frequently got an error saying that I needed to include a .name_repair argument in my as_tibble function.*

```{r}
#Copying the function that returns a single year into the same code chunk as the new function so that it all can be run together smoothly
return_date <- function(year = 2024){
  rds <- paste0(year, '_variables.rds')
  variables <- readRDS(rds)
  return(variables)
}

#Function that returns multiple years
combine_years <- function(return_date, survey_years){
  results <- list()
   for(i in seq_along(survey_years)){
     results[[i]] <- return_date(survey_years[i])
   }
  tibble_results <- as_tibble(results, .name_repair = 'universal')
  return(tibble_results)
}

#Testing it out, using str() as I expect that this will yield a very large tibble
test_combine <- combine_years(return_date, c(2021, 2022, 2023, 2024))
str(test_combine)
```

# PART TWO: WRITING A GENERIC FUNCTION FOR SUMMARIZING

## NUMERIC SUMMARIES

```{r}

```

## GRAPHICAL SUMMARIES

```{r}

```
