---
title: "project_1_code"
format: html
editor: visual
---

# PART 1: DATA PROCESSING

## PARSING VARIABLE VALUES

```{r}
#Reading in packages to current session
library(httr)
library(jsonlite)
library(dplyr)
library(tidyverse)
```

*Writing functions to parse variable data.*

```{r}
#Writing functions to get urls and parse them. Note that the parsed_data objects on their own are very large and take a long time to load. It's easier to use str() to look at them.
list_vars_21 <- function(api_url1){
  api_url1 <- 'https://api.census.gov/data/2021/acs/acs1/pums/variables.json'
  data_url1 <- httr::GET(api_url1)
  parsed_data1 <- fromJSON(rawToChar(data_url1$content))
}
variables_2021 <- list_vars_21(api_url1)

list_vars_22 <- function(api_url2){
  api_url2 <- 'https://api.census.gov/data/2022/acs/acs1/pums/variables.json'
  data_url2 <- httr::GET(api_url2)
  parsed_data2 <- fromJSON(rawToChar(data_url2$content))
}
variables_2022 <- list_vars_22(api_url2)

list_vars_23 <- function(api_url3){
  api_url3 <- 'https://api.census.gov/data/2023/acs/acs1/pums/variables.json'
  data_url3 <- httr::GET(api_url3)
  parsed_data3 <- fromJSON(rawToChar(data_url3$content))
}
variables_2023 <- list_vars_23(api_url3)

list_vars_24 <- function(api_url4){
  api_url4 <- 'https://api.census.gov/data/2024/acs/acs1/pums/variables.json'
  data_url4 <- httr::GET(api_url4)
  parsed_data4 <- fromJSON(rawToChar(data_url4$content))
}
variables_2024 <- list_vars_24(api_url4)
```

*Putting variable lists into .rds files.*

```{r}
#Previously attempted this by putting everything into one data frame, but it ended up recycling parts due to uneven lengths.
saveRDS(variables_2021, file = '2021_variables.rds')
saveRDS(variables_2022, file = '2022_variables.rds')
saveRDS(variables_2023, file = '2023_variables.rds')
saveRDS(variables_2024, file = '2024_variables.rds')
```

*Testing one of the functions to see if it works. The function works with both the link to the variables and the api_url object.*

```{r}
#Must include variables column when retrieving a specific element, otherwise it will return a NULL.
test_24 <- list_vars_24(api_url4)
test_24$variables$SEX
```

## PARSING API DATA

*Getting URL, writing helper function, and parsing data.*

```{r}

```

```{r}

```

*Writing function that allows user to change year of survey.*

```{r}

```

*Writing functions to specify numeric variables to be returned*.

```{r}


```

*Writing functions to specify categorical variables to be returned.*

```{r}

```

*Writing function to specify geography level.*

```{r}

```

*Writing function to specify region, division, or state to subset data on.*

```{r}

```

## COMBINING YEARS

*Writing function to specify multiple years of survey data.*

```{r}

```

# PART TWO: WRITING A GENERIC FUNCTION FOR SUMMARIZING

## NUMERIC SUMMARIES

```{r}

```

## GRAPHICAL SUMMARIES

```{r}

```
