---
title: "project_1_code"
format: html
editor: visual
---

# PART 1: DATA PROCESSING

## PARSING VARIABLE VALUES

```{r}
#Reading in packages to current session
library(httr)
library(jsonlite)
library(dplyr)
library(tidyverse)
```

*Writing functions to parse variable data.*

```{r}
#Writing functions to get urls and parse them. Note that the parsed_data objects on their own are very large and take a long time to load. It's easier to use str() to look at them.
list_vars_21 <- function(api_url1){
  api_url1 <- 'https://api.census.gov/data/2021/acs/acs1/pums/variables.json'
  data_url1 <- httr::GET(api_url1)
  parsed_data1 <- fromJSON(rawToChar(data_url1$content))
}
variables_2021 <- list_vars_21(api_url1)

list_vars_22 <- function(api_url2){
  api_url2 <- 'https://api.census.gov/data/2022/acs/acs1/pums/variables.json'
  data_url2 <- httr::GET(api_url2)
  parsed_data2 <- fromJSON(rawToChar(data_url2$content))
}
variables_2022 <- list_vars_22(api_url2)

list_vars_23 <- function(api_url3){
  api_url3 <- 'https://api.census.gov/data/2023/acs/acs1/pums/variables.json'
  data_url3 <- httr::GET(api_url3)
  parsed_data3 <- fromJSON(rawToChar(data_url3$content))
}
variables_2023 <- list_vars_23(api_url3)

list_vars_24 <- function(api_url4){
  api_url4 <- 'https://api.census.gov/data/2024/acs/acs1/pums/variables.json'
  data_url4 <- httr::GET(api_url4)
  parsed_data4 <- fromJSON(rawToChar(data_url4$content))
}
variables_2024 <- list_vars_24(api_url4)
```

*Putting variable lists into .Rmd file*.

```{r}
saveRDS(variables_2021, variables_2022, variables_2023, variables_2024, file = 'variable_lists.Rmd')
```

*Testing one of the functions to see if it works. The function works with both the link to the variables and the api_url object.*

```{r}
#Must include variables column when retrieving a specific element, otherwise it will return a NULL.
test_24 <- list_vars_24(api_url4)
test_24$variables$SEX
```

## PARSING API DATA

*Getting URL, writing helper function, and parsing data.*

```{r}
library(lubridate)
#Getting 2021 survey
pums_data_21 <- httr::GET('api.census.gov/data/2021/acs/acs1/pums?get=AGEP,GRPIP,JWAP,JWDP,JWMNP,PWGTP,FER,HHL,SCH,SCHL,SEX,ST,REGION,DIVISION&key=97196a6e47d04457a6ffdf399f45a446')

#Using function to make it into a tibble
helper_21 <- function(pums_data_21){
  parsed_data <- rawToChar(pums_data_21[content])
  tibble_data <- as.tibble(parsed_data[data])
}

#Creating named object from tibble
parsed_tibble_21 <- helper_21(pums_data_21)

#Getting 2022 survey
pums_data_22 <- httr::GET('api.census.gov/data/2022/acs/acs1/pums?get=AGEP,GRPIP,JWAP,JWDP,JWMNP,PWGTP,FER,HHL,SCH,SCHL,SEX,ST,REGION,DIVISION&key=97196a6e47d04457a6ffdf399f45a446')

#Using function to make it into a tibble
helper_22 <- function(pums_data_22){
  tibble_data2 <- as.tibble(pums_data_22$data)
  raw_data2 <- as.raw(tibble_data2)
  parsed_data2 <- rawToChar(raw_data2$content)
}

#Creating named object from tibble
parsed_tibble_22 <- helper_22(pums_data_22)

#Getting 2023 survey
pums_data_23 <- httr::GET('api.census.gov/data/2023/acs/acs1/pums?get=AGEP,GRPIP,JWAP,JWDP,JWMNP,PWGTP,FER,HHL,SCH,SCHL,SEX,ST,REGION,DIVISION&key=97196a6e47d04457a6ffdf399f45a446')

#Using function to make it into a tibble
helper_23 <- function(pums_data_23){
  tibble_data3 <- as.tibble(pums_data_23$data)
  raw_data3 <- as.raw(tibble_data3)
  parsed_data3 <- rawToChar(raw_data3$content)
}

#Creating named object from tibble
parsed_tibble_23 <- helper_23(pums_data_23)

#Getting 2024 survey
pums_data_24 <- httr::GET('api.census.gov/data/2024/acs/acs1/pums?get=AGEP,GRPIP,JWAP,JWDP,JWMNP,PWGTP,FER,HHL,SCH,SCHL,SEX,STATE,REGION,DIVISION&key=97196a6e47d04457a6ffdf399f45a446')

#Using function to make it into a tibble
helper_24 <- function(pums_data_24){
  tibble_data4 <- as.tibble(pums_data_24$data)
  raw_data4 <- as.raw(tibble_data4)
  parsed_data4 <- rawToChar(raw_data4$content)
}

#Creating named object from tibble
parsed_tibble_24 <- helper_24(pums_data_24)

#Checking structure of tibbles to make sure that specific elements are being called correctly.
str(parsed_tibble_21)
str(parsed_tibble_22)
str(parsed_tibble_23)
str(parsed_tibble_24)
```

```{r}

```

*Writing function that allows user to change year of survey.*

```{r}
#Writing function to return year of survey
return_date <- function('name of parsed variables'){
  if(list_vars_21){
    year_of_survey <- '2021'
    return(year_of_survey)
  } else if(list_vars_22){
    year_of_survey <- '2022'
    return(year_of_survey)
  } else if(list_vars_23){
    year_of_survey <- '2023'
    return(year_of_survey)
  } else if(list_vars_24){
    year_of_survey <- '2024'
    return(year_of_survey)
  }
}
```

*Writing functions to specify numeric variables to be returned*.

```{r}
#Planning out function to change variables to numeric
change_num_variable <- function('name of parsed data$variables$target var'){
  if(parsed_data$variables$AGEP, parsed_data$variables$GASP, parsed_data$variables$GRPIP, parsed_data$variables$JWMNP){
    as.numeric(parsed_data$variables$AGEP, parsed_data$variables$GASP, parsed_data$variables$GRPIP, parsed_data$variables$JWMNP)
}

#Planning out function to change variables to time
change_time_variable <- function('name of data'){
  time_JWAP_start <- 'use helper data$variables$JWAP'
  time_JWAP_end <- 'use helper data$variables$JWAP'
  time_JWDP_start <- 'use helper data$variables$JWDP'
  time_JWAP_end <- 'use helper data$variables$JWDP'
  hms_JWAP_start <- hms(time_JWAP_start)
  hms_JWAP_end <- hms(time_JWAP_end)
  hms_JWDP_start <- hms(time_JWDP_start)
  hms_JWDP_end <- hms(time_JWDP_end)
  average_JWAP <- mean(c(hms_JWAP_start, hms_JWAP_end))
  average_JWDP <- mean(c(hms_JWDP_start, hms_JWDP_end))
  midpoint_JWAP <- format(as_datetime(average_JWAP), '%H:%M:%S')
  midpoint_JWDP <- format(as_datetime(average_JWDP), '%H:%M:%S')
  print(midpoint_JWAP, midpoint_JWDP)
}

#Possibly a better helper function?
change_time_variable <- function('inputs'){
  time_JWAP <- 'use parsed tibble'
  time_JWDP <- 'use parsed tibble'
  JWAP_midpoint <- lapply(X = time_JWAP, FUN = mean.difftime, na.rm = TRUE)
  JWDP_midpoint <- lapply(X = time_JWDP, FUN = mean.difftime, na.rm = TRUE)
  print(JWAP_midpoint, JWDP_midpoint)
}

```

*Writing functions to specify categorical variables to be returned.*

```{r}
categorical_helper <- function('input'){
  category_lists <- readRDS('file')
}

#Writing function to return categorical variables
return_cat_vars <- function('input'){
  cat_helper <- function('input'){
    category_lists <- readRDS('file')
  }
  as.factor('list the variables here')
  
}
```

*Writing function to specify geography level.*

```{r}

```

*Writing function to specify region, division, or state to subset data on.*

```{r}
#Specifying region, division, or state
region_division_state <- function('input'){
  
}
```

## COMBINING YEARS

*Writing function to specify multiple years of survey data.*

```{r}

```

# PART TWO: WRITING A GENERIC FUNCTION FOR SUMMARIZING

## NUMERIC SUMMARIES

```{r}
#Writing function for numerical summaries
summary.census <- function(tibble, numeric_vector, weight_vector){
  tibble <- 'put tibble here'
  numeric_vector <- 'put numeric vector here'
  weight_vector <- 'put weight vector here'
  sample_mean <- sum(numeric_vector*weight_vector)/sum(weight_vector)
  sample_std_dev <- sqrt(sum(numeric_vectorˆ2*weight_vector)/sum(weight_vector)-sample_meanˆ2)
  return(list(sample_mean, sample_std_dev))
}
```

## GRAPHICAL SUMMARIES

```{r}
#Writing function for graphical summaries
box_plot <- function(cat, num){
  return(plot('name', aes(x = cat, y = num, fill = cat, weight = PWGTP)
}
```
