---
title: "project_1_code"
format: html
editor: visual
---

### EMMA BERNINGER

### TURNED IN: 10/5/2O26

### PROJECT 1 FINAL CODE

# PART 1: DATA PROCESSING

## PARSING VARIABLE VALUES

```{r}
#Reading in packages to current session
library(httr)
library(jsonlite)
library(dplyr)
library(tidyverse)
library(lubridate)
library(stringr)
```

*Writing functions to parse variable data.*

```{r}
#Writing functions to get urls and parse them. Note that the parsed_data objects on their own are very large and take a long time to load. It's easier to use str() to look at them.
list_vars_21 <- function(api_url1){
  api_url1 <- 'https://api.census.gov/data/2021/acs/acs1/pums/variables.json'
  data_url1 <- httr::GET(api_url1)
  parsed_data1 <- fromJSON(rawToChar(data_url1$content))
}
variables_2021 <- list_vars_21(api_url1)

list_vars_22 <- function(api_url2){
  api_url2 <- 'https://api.census.gov/data/2022/acs/acs1/pums/variables.json'
  data_url2 <- httr::GET(api_url2)
  parsed_data2 <- fromJSON(rawToChar(data_url2$content))
}
variables_2022 <- list_vars_22(api_url2)

list_vars_23 <- function(api_url3){
  api_url3 <- 'https://api.census.gov/data/2023/acs/acs1/pums/variables.json'
  data_url3 <- httr::GET(api_url3)
  parsed_data3 <- fromJSON(rawToChar(data_url3$content))
}
variables_2023 <- list_vars_23(api_url3)

list_vars_24 <- function(api_url4){
  api_url4 <- 'https://api.census.gov/data/2024/acs/acs1/pums/variables.json'
  data_url4 <- httr::GET(api_url4)
  parsed_data4 <- fromJSON(rawToChar(data_url4$content))
}
variables_2024 <- list_vars_24(api_url4)
```

*Putting variable lists into .rds files.*

*I had previously tried putting everything into one data frame, but it ended up recycling parts due to uneven lengths. Having them in four separate files seems to work fine.*

```{r}
#Previously attempted this by putting everything into one data frame, but it ended up recycling parts due to uneven lengths.
saveRDS(variables_2021, file = '2021_variables.rds')
saveRDS(variables_2022, file = '2022_variables.rds')
saveRDS(variables_2023, file = '2023_variables.rds')
saveRDS(variables_2024, file = '2024_variables.rds')
```

*Testing the functions to see if they work. The functions work with both the link to the variables and the api_url object.*

```{r}
#Must include variables column when retrieving a specific element, otherwise it will return a NULL.
test_21 <- list_vars_21(api_url1)
test_21$variables$ST

test_22 <- list_vars_22(api_url2)
test_22$variables$AGEP

test_23 <- list_vars_23(api_url3)
test_23$variables$GRPIP

test_24 <- list_vars_24(api_url4)
test_24$variables$SEX
```

## PARSING API DATA

*Getting URL, writing helper function, and parsing data.*

*These take a long time to run, querying the API directly through a link takes a long time for my computer to process.*

*These presented my first problems with this project, I was getting errors about using both \$ and \[\] within my function to call certain objects. In the as_tibble function, I had been calling the object parsed_data\$data. I had seen in the reading and remembered from previous homework that that's something you should do when reading in API data. I found out that this was the issue when I accidentally wrote a function similar to this one 'wrong' and it ended up working perfectly fine. The part that was different was the lack of '\$data', as I had forgotten it.*

```{r}
#Getting 2021 survey
pums_data_21 <- httr::GET('api.census.gov/data/2021/acs/acs1/pums?get=AGEP,GRPIP,JWAP,JWDP,JWMNP,PWGTP,FER,HHL,SCH,SCHL,SEX,ST,REGION,DIVISION&key=416ac733e884381d5fc80215dc879af0edb1dd8b')

#Using function to make it into a tibble
helper_21 <- function(pums_data_21){
  parsed_data <- fromJSON(rawToChar(pums_data_21$content))
  tibble_data <- as_tibble(parsed_data)
}

#Creating named object from tibble and checking structure
parsed_tibble_21 <- helper_21(pums_data_21)
str(parsed_tibble_21)
```

```{r}
#Getting 2022 survey
pums_data_22 <- httr::GET('api.census.gov/data/2022/acs/acs1/pums?get=AGEP,GRPIP,JWAP,JWDP,JWMNP,PWGTP,FER,HHL,SCH,SCHL,SEX,ST,REGION,DIVISION&key=416ac733e884381d5fc80215dc879af0edb1dd8b')

#Using function to make it into a tibble
helper_22 <- function(pums_data_22){
  parsed_data2 <- fromJSON(rawToChar(pums_data_22$content))
  tibble_data2 <- as_tibble(parsed_data2)
}

#Creating named object from tibble and checking structure
parsed_tibble_22 <- helper_22(pums_data_22)
str(parsed_tibble_22)
```

```{r}
#Getting 2023 survey
pums_data_23 <- httr::GET('api.census.gov/data/2023/acs/acs1/pums?get=AGEP,GRPIP,JWAP,JWDP,JWMNP,PWGTP,FER,HHL,SCH,SCHL,SEX,STATE,REGION,DIVISION&key=416ac733e884381d5fc80215dc879af0edb1dd8b')

#Using function to make it into a tibble
helper_23 <- function(pums_data_23){
  parsed_data3 <- fromJSON(rawToChar(pums_data_23$content))
  tibble_data3 <- as_tibble(parsed_data3)
}

#Creating named object from tibble and checking structure
parsed_tibble_23 <- helper_23(pums_data_23)
str(parsed_tibble_23)
```

```{r}
#Getting 2024 survey
pums_data_24 <- httr::GET('api.census.gov/data/2024/acs/acs1/pums?get=AGEP,GRPIP,JWAP,JWDP,JWMNP,PWGTP,FER,HHL,SCH,SCHL,SEX,STATE,REGION,DIVISION&key=416ac733e884381d5fc80215dc879af0edb1dd8b')

#Using function to make it into a tibble
helper_24 <- function(pums_data_24){
  parsed_data4 <- fromJSON(rawToChar(pums_data_24$content))
  tibble_data4 <- as_tibble(parsed_data4)
}

#Creating named object from tibble and checking structure
parsed_tibble_24 <- helper_24(pums_data_24)
str(parsed_tibble_24)
```

*Writing function that allows user to change year of survey.*

*I previously was using if/else logic to write these functions, in the style below:*

my_function \<- function(year){

year \<- c(2021, 2022, 2023, 2024)

if(2021){

Code to be run

}else if(2022){

Code to be run

} and so on.

*These functions turned out to be extremely long. I tried paste0 to make selecting the year of the data easier and it worked, so I was able to shorten them significantly.*

```{r}
#Writing function to return year of survey
return_date <- function(year = 2024){
  rds <- paste0(year, '_variables.rds')
  variables <- readRDS(rds)
  return(variables)
}

#Testing this function
test_date <- return_date(2021)
str(test_date)
```

*Writing functions to change numeric variables from characters to numeric and then return them.*

*With this function, I had the most amount of trouble with the type of objects the variables were. I was continuously getting errors and warnings about all sorts of things, particularly when trying to convert them to numeric variables. Finally, making them into vectors worked.*

```{r}
#Writing function that returns numeric variables, also changes them from characters to numeric
return_num_vars <- function(year, target_var = 'AGEP'){
  rds <- paste0(year, "_variables.rds")
  var_list <- readRDS(rds)
  PWGTP = c(var_list$variables$PWGTP$values$range$min, var_list$variables$PWGTP$values$range$max)
  AGEP = c(var_list$variables$AGEP$values$range$min, var_list$variables$AGEP$values$range$max)
  GASP = c(var_list$variables$GASP$values$range$min, var_list$variables$GASP$values$range$max)
  GRPIP = c(var_list$variables$GRPIP$values$range$min, var_list$variables$GRPIP$values$range$max)
  JWMNP = c(var_list$variables$JWMNP$values$range$min, var_list$variables$JWMNP$values$range$max)
  full_list <- list(AGEP = AGEP, GASP = GASP, GRPIP = GRPIP, JWMNP = JWMNP, PWGTP = PWGTP)
  num_list <- lapply(full_list, as.numeric, na.rm = TRUE)
  return(list(num_list['PWGTP'], num_list[target_var]))
}

test_num <- return_num_vars(year = 2021, target_var = 'GASP')
test_num
```

*Writing a function to turn the time data into numeric time data.*

*This was definitely the part that gave me the most trouble. There was a point where the function I was trying to write would cause RStudio to shut down and become completely unresponsive. I also decided to separate this into two different functions because I imagine the combined one would be very long.*

```{r}
#Writing code to return midpoint of JWAP
return_JWAP <- function(year){
variables <- paste0(year, '_variables.rds')
variables_list <- readRDS(variables)
variables_df <- data.frame(
  data = unlist(variables_list$variables$JWAP$values$item),
  stringsAsFactors = FALSE
)

to_minutes <- function(time_str, am_pm){
  parts <- str_split(time_str, ':', simplify = TRUE)
  hours <- as.numeric(parts[, 1])
  minutes <- as.numeric(parts[, 2])
  
  is_pm <- str_detect(am_pm, 'p')
  
  hours <- case_when(
    is_pm & hours != 12 ~ hours + 12,
    !is_pm & hours == 12 ~ 0,
    TRUE                 ~ hours
  )
  return(hours * 60 + minutes)
}
processed <- variables_df |>
  mutate(
    start_time = str_extract(data, "^\\d{1,2}:\\d{2}"),
    start_ampm = str_extract(data, "^\\d{1,2}:\\d{2}\\s*[ap]\\.m\\."),
    end_time = str_extract(data, "\\d{1,2}:\\d{2}\\s*[ap]\\.m\\.$") %>% str_extract("^\\d{1,2}:\\d{2}"),
    end_ampm = str_extract(data, "\\d{1,2}:\\d{2}\\s*[ap]\\.m\\.$"),
    start_mins = to_minutes(start_time, start_ampm),
    end_mins = to_minutes(end_time, end_ampm),
    end_mins = if_else(end_mins < start_mins, end_mins + 1440, end_mins),
    midpoint_mins = (start_mins + end_mins) / 2,
    midpoint_mins = midpoint_mins %% 1440,
    midpoint_time = sprintf("%02d:%02d", floor(midpoint_mins / 60), round(midpoint_mins %% 60))
  ) |>
  select(midpoint_time)
print(processed)
}


#Writing code to return midpoint of JWDP
return_JWDP <- function(year){
  variables <- paste0(year, '_variables.rds')
variables_list <- readRDS(variables)
variables_df <- data.frame(
  data = unlist(variables_list$variables$JWDP$values$item),
  stringsAsFactors = FALSE
)

to_minutes <- function(time_str, am_pm){
  parts <- str_split(time_str, ':', simplify = TRUE)
  hours <- as.numeric(parts[, 1])
  minutes <- as.numeric(parts[, 2])
  
  is_pm <- str_detect(am_pm, 'p')
  
  hours <- case_when(
    is_pm & hours != 12 ~ hours + 12,
    !is_pm & hours == 12 ~ 0,
    TRUE                 ~ hours
  )
  return(hours * 60 + minutes)
}
processed <- variables_df |>
  mutate(
    start_time = str_extract(data, "^\\d{1,2}:\\d{2}"),
    start_ampm = str_extract(data, "^\\d{1,2}:\\d{2}\\s*[ap]\\.m\\."),
    end_time = str_extract(data, "\\d{1,2}:\\d{2}\\s*[ap]\\.m\\.$") %>% str_extract("^\\d{1,2}:\\d{2}"),
    end_ampm = str_extract(data, "\\d{1,2}:\\d{2}\\s*[ap]\\.m\\.$"),
    start_mins = to_minutes(start_time, start_ampm),
    end_mins = to_minutes(end_time, end_ampm),
    end_mins = if_else(end_mins < start_mins, end_mins + 1440, end_mins),
    midpoint_mins = (start_mins + end_mins) / 2,
    midpoint_mins = midpoint_mins %% 1440,
    midpoint_time = sprintf("%02d:%02d", floor(midpoint_mins / 60), round(midpoint_mins %% 60))
  ) |>
  select(midpoint_time)
print(processed)
}

#Examples of how to use this function
JWAP_test <- return_JWAP(2021)
JWAP_test
JWDP_test <- return_JWDP(2021)
JWDP_test
```

*Writing functions to specify categorical variables to be returned.*

*Writing these didn't give me too much trouble outside of figuring out how to change the variables into factors. I had been previously been using as.factor(), which had not been working.*

```{r}
#Writing function
return_cat_vars <- function(year, target_var = 'SEX'){
rds <- paste0(year, '_variables.rds')
var_list <- readRDS(rds)
FER = factor(unlist(var_list$variables$FER))
HHL = factor(unlist(var_list$variables$HHL))
SCH = factor(unlist(var_list$variables$SCH))
SCHL = factor(unlist(var_list$variables$SCHL))
SEX = factor(unlist(var_list$variables$SEX))
var_list <- list(FER = FER, HHL = HHL, SCH = SCH, SCHL = SCHL, SEX = SEX)
return(var_list[[target_var]])
}

#Example of how to use this function
test_cat_vars <- return_cat_vars(2023, 'FER')
test_cat_vars
```

*Writing function to specify geography level.*

*I didn't have much trouble with this one. I decided to split them into two functions to account for the STATE and ST variables.*

```{r}
#This one is meant for the years 2021 and 2022
geography_level <- function(year, target_var = 'REGION', all_values = FALSE){
rds <- paste0(year, '_variables.rds')
vars_list <- readRDS(rds)
REGION = vars_list$variables$REGION
ST = vars_list$variables$ST
DIVISION = vars_list$variables$DIVISION
geo_list <- list(REGION = REGION, ST = ST, DIVISION = DIVISION)
if(all_values){
  return('ERROR')
}else{
  return(geo_list[[target_var]])
}
}

#This one is meant for the years 2023 and 2024
geography_level_2 <- function(year, target_var = 'REGION', all_values = FALSE){
rds <- paste0(year, '_variables.rds')
vars_list <- readRDS(rds)
REGION = vars_list$variables$REGION
STATE = vars_list$variables$STATE
DIVISION = vars_list$variables$DIVISION
geo_list <- list(REGION = REGION, STATE = STATE, DIVISION = DIVISION)
if(all_values){
  return('ERROR')
}else{
  return(geo_list[[target_var]])
}
}

#Examples of how to use these functions, including one that returns 'error'.
test_geo <- geography_level(2021, target_var = 'DIVISION')
test_geo
test_geo_2 <- geography_level_2(2023, all_values = TRUE)
test_geo_2
```

*Writing functions to specify region, division, or state to subset data on.*

*This group was very easy to put together.*

```{r}
#Writing the region funcition
subset_region <- function(year, target_var = 'Northeast', all_values = FALSE){
rds <- paste0(year, '_variables.rds')
vars_list <- readRDS(rds)
Northeast = vars_list$variables$REGION$values$item$`1`
Puerto_Rico = vars_list$variables$REGION$values$item$`9`
South = vars_list$variables$REGION$values$item$`3`
West = vars_list$variables$REGION$values$item$`4`
Midwest = vars_list$variables$REGION$values$item$`2`
region_list <- list(Northeast = Northeast, Puerto_Rico = Puerto_Rico, South = South, West = West, Midwest = Midwest)
if(all_values){
  return(region_list)
}else{
  return(region_list[[target_var]])
}
}
#Example of how to use this function
test_geo <- subset_region(2023, target_var = 'South', all_values = FALSE)
test_geo
```

```{r}
#Writing the state function

#This one should be used for the 2021 and 2022 data
subset_st <- function(year, target_var = 'North_Carolina', all_values = FALSE){
rds <- paste0(year, '_variables.rds')
vars_list <- readRDS(rds)
Tennessee = vars_list$variables$ST$values$item$`47`
Arizona = vars_list$variables$ST$values$item$`04`
Arkansas = vars_list$variables$ST$values$item$`05`
Utah = vars_list$variables$ST$values$item$`49`
Connecticut = vars_list$variables$ST$values$item$`09`
Vermont = vars_list$variables$ST$values$item$`50`
Mississippi = vars_list$variables$ST$values$item$`28`
North_Carolina = vars_list$variables$ST$values$item$`37`
Pennsylvania = vars_list$variables$ST$values$item$`42`
South_Carolina = vars_list$variables$ST$values$item$`45`
Virginia = vars_list$variables$ST$values$item$`51`
Hawaii = vars_list$variables$ST$values$item$`15`
Massachusetts = vars_list$variables$ST$values$item$`25`
Puerto_Rico = vars_list$variables$ST$values$item$`72`
Montana = vars_list$variables$ST$values$item$`30`
New_Hampshire = vars_list$variables$ST$values$item$`33`
New_Mexico = vars_list$variables$ST$values$item$`35`
Florida = vars_list$variables$ST$values$item$`12`
Texas = vars_list$variables$ST$values$item$`48`
Washington = vars_list$variables$ST$values$item$`53`
Georgia = vars_list$variables$ST$values$item$`13`
Indiana = vars_list$variables$ST$values$item$`18`
Kansas = vars_list$variables$ST$values$item$`20`
Michigan = vars_list$variables$ST$values$item$`26`
Nevada = vars_list$variables$ST$values$item$`32`
West_Virginia = vars_list$variables$ST$values$item$`54`
Wisconsin = vars_list$variables$ST$values$item$`55`
Minnesota = vars_list$variables$ST$values$item$`27`
Idaho = vars_list$variables$ST$values$item$`16`
Iowa = vars_list$variables$ST$values$item$`19`
Louisiana = vars_list$variables$ST$values$item$`22`
North_Dakota = vars_list$variables$ST$values$item$`38`
Alabama = vars_list$variables$ST$values$item$`01`
Alaska = vars_list$variables$ST$values$item$`02`
California = vars_list$variables$ST$values$item$`06`
Colorado = vars_list$variables$ST$values$item$`08`
District_of_Columbia = vars_list$variables$ST$values$item$`11`
Maine = vars_list$variables$ST$values$item$`23`
New_Jersey = vars_list$variables$ST$values$item$`34`
Ohio = vars_list$variables$ST$values$item$`39`
Delaware = vars_list$variables$ST$values$item$`10`
Wyoming = vars_list$variables$ST$values$item$`56`
Kentucky = vars_list$variables$ST$values$item$`21`
Maryland = vars_list$variables$ST$values$item$`24`
Rhode_Island = vars_list$variables$ST$values$item$`44`
South_Dakota = vars_list$variables$ST$values$item$`46`
Illinois = vars_list$variables$ST$values$item$`17`
Missouri = vars_list$variables$ST$values$item$`29`
Nebraska = vars_list$variables$ST$values$item$`31`
New_York = vars_list$variables$ST$values$item$`36`
Oklahoma = vars_list$variables$ST$values$item$`40`
Oregon = vars_list$variables$ST$values$item$`41`
st_list <- list(Tennessee = Tennessee, Arizona = Arizona, Arkansas = Arkansas, Utah = Utah, Connecticut = Connecticut, Vermont = Vermont, Mississippi = Mississippi, North_Carolina = North_Carolina, Pennsylvania = Pennsylvania, South_Carolina = South_Carolina, Virginia = Virginia, Hawaii = Hawaii, Massachusetts = Massachusetts, Puerto_Rico = Puerto_Rico, Montana = Montana, New_Hampshire = New_Hampshire, New_Mexico = New_Mexico, Florida = Florida, Texas = Texas, Washington = Washington, Georgia = Georgia, Indiana = Indiana, Kansas = Kansas, Michigan = Michigan, Nevada = Nevada, West_Virginia = West_Virginia, Wisconsin = Wisconsin, Minnesota = Minnesota, Idaho = Idaho, Iowa = Iowa, Louisiana = Louisiana, North_Dakota = North_Dakota, Alabama = Alabama, Alaska = Alaska, California = California, Colorado = Colorado, District_of_Columbia = District_of_Columbia, Maine = Maine, New_Jersey = New_Jersey, Ohio = Ohio, Delaware = Delaware, Wyoming = Wyoming, Kentucky = Kentucky, Maryland = Maryland, Rhode_Island = Rhode_Island, South_Dakota = South_Dakota, Illinois = Illinois, Missouri = Missouri, Nebraska = Nebraska, New_York = New_York, Oklahoma = Oklahoma, Oregon = Oregon)

if(all_values){
  return(st_list)
}else{
  return(st_list[[target_var]])
}
}

#This one should be used for the 2023 and 2024 data
subset_state <- function(year, target_var = 'North_Carolina', all_values = FALSE){
rds <- paste0(year, '_variables.rds')
vars_list <- readRDS(rds)
Tennessee = vars_list$variables$STATE$values$item$`47`
Arizona = vars_list$variables$STATE$values$item$`04`
Arkansas = vars_list$variables$STATE$values$item$`05`
Utah = vars_list$variables$STATE$values$item$`49`
Connecticut = vars_list$variables$STATE$values$item$`09`
Vermont = vars_list$variables$STATE$values$item$`50`
Mississippi = vars_list$variables$STATE$values$item$`28`
North_Carolina = vars_list$variables$STATE$values$item$`37`
Pennsylvania = vars_list$variables$STATE$values$item$`42`
South_Carolina = vars_list$variables$STATE$values$item$`45`
Virginia = vars_list$variables$STATE$values$item$`51`
Hawaii = vars_list$variables$STATE$values$item$`15`
Massachusetts = vars_list$variables$STATE$values$item$`25`
Puerto_Rico = vars_list$variables$STATE$values$item$`72`
Montana = vars_list$variables$STATE$values$item$`30`
New_Hampshire = vars_list$variables$STATE$values$item$`33`
New_Mexico = vars_list$variables$STATE$values$item$`35`
Florida = vars_list$variables$STATE$values$item$`12`
Texas = vars_list$variables$STATE$values$item$`48`
Washington = vars_list$variables$STATE$values$item$`53`
Georgia = vars_list$variables$STATE$values$item$`13`
Indiana = vars_list$variables$STATE$values$item$`18`
Kansas = vars_list$variables$STATE$values$item$`20`
Michigan = vars_list$variables$STATE$values$item$`26`
Nevada = vars_list$variables$STATE$values$item$`32`
West_Virginia = vars_list$variables$STATE$values$item$`54`
Wisconsin = vars_list$variables$STATE$values$item$`55`
Minnesota = vars_list$variables$STATE$values$item$`27`
Idaho = vars_list$variables$STATE$values$item$`16`
Iowa = vars_list$variables$STATE$values$item$`19`
Louisiana = vars_list$variables$STATE$values$item$`22`
North_Dakota = vars_list$variables$STATE$values$item$`38`
Alabama = vars_list$variables$STATE$values$item$`01`
Alaska = vars_list$variables$STATE$values$item$`02`
California = vars_list$variables$STATE$values$item$`06`
Colorado = vars_list$variables$STATE$values$item$`08`
District_of_Columbia = vars_list$variables$STATE$values$item$`11`
Maine = vars_list$variables$STATE$values$item$`23`
New_Jersey = vars_list$variables$STATE$values$item$`34`
Ohio = vars_list$variables$STATE$values$item$`39`
Delaware = vars_list$variables$STATE$values$item$`10`
Wyoming = vars_list$variables$STATE$values$item$`56`
Kentucky = vars_list$variables$STATE$values$item$`21`
Maryland = vars_list$variables$STATE$values$item$`24`
Rhode_Island = vars_list$variables$STATE$values$item$`44`
South_Dakota = vars_list$variables$STATE$values$item$`46`
Illinois = vars_list$variables$STATE$values$item$`17`
Missouri = vars_list$variables$STATE$values$item$`29`
Nebraska = vars_list$variables$STATE$values$item$`31`
New_York = vars_list$variables$STATE$values$item$`36`
Oklahoma = vars_list$variables$STATE$values$item$`40`
Oregon = vars_list$variables$STATE$values$item$`41`
state_list <- list(Tennessee = Tennessee, Arizona = Arizona, Arkansas = Arkansas, Utah = Utah, Connecticut = Connecticut, Vermont = Vermont, Mississippi = Mississippi, North_Carolina = North_Carolina, Pennsylvania = Pennsylvania, South_Carolina = South_Carolina, Virginia = Virginia, Hawaii = Hawaii, Massachusetts = Massachusetts, Puerto_Rico = Puerto_Rico, Montana = Montana, New_Hampshire = New_Hampshire, New_Mexico = New_Mexico, Florida = Florida, Texas = Texas, Washington = Washington, Georgia = Georgia, Indiana = Indiana, Kansas = Kansas, Michigan = Michigan, Nevada = Nevada, West_Virginia = West_Virginia, Wisconsin = Wisconsin, Minnesota = Minnesota, Idaho = Idaho, Iowa = Iowa, Louisiana = Louisiana, North_Dakota = North_Dakota, Alabama = Alabama, Alaska = Alaska, California = California, Colorado = Colorado, District_of_Columbia = District_of_Columbia, Maine = Maine, New_Jersey = New_Jersey, Ohio = Ohio, Delaware = Delaware, Wyoming = Wyoming, Kentucky = Kentucky, Maryland = Maryland, Rhode_Island = Rhode_Island, South_Dakota = South_Dakota, Illinois = Illinois, Missouri = Missouri, Nebraska = Nebraska, New_York = New_York, Oklahoma = Oklahoma, Oregon = Oregon)
if(all_values){
  return(state_list)
}else{
  return(state_list[[target_var]])
}
}
#Examples of how to use these functions
test_st <- subset_st(2021, 'Alabama', all_values = FALSE)
test_st

test_state <- subset_state(2024, 'Tennessee', all_values = FALSE)
test_state
```

```{r}
#Writing the division function
subset_division <- function(year, target_var = 'Puerto_Rico', all_values = FALSE){
rds <- paste0(year, '_variables.rds')
vars_list <- readRDS(rds)
West_South_Central_South_Region = vars_list$variables$DIVISION$values$item$`7`
Middle_Atlantic_Northeast_Region = vars_list$variables$DIVISION$values$item$`2`
West_North_Central_Midwest_Region = vars_list$variables$DIVISION$values$item$`4`
East_South_Central_South_Region = vars_list$variables$DIVISION$values$item$`6`
East_North_Central_Midwest_Region = vars_list$variables$DIVISION$values$item$`3`
Pacific_West_Region = vars_list$variables$DIVISION$values$item$`9`
Puerto_Rico = vars_list$variables$DIVISION$values$item$`0`
South_Atlantic_South_Region = vars_list$variables$DIVISION$values$item$`5`
Mountain_West_Region = vars_list$variables$DIVISION$values$item$`8`
New_England_Northeast_Region = vars_list$variables$DIVISION$values$item$`1`
div_list <- list(West_South_Central_South_Region = West_South_Central_South_Region, Middle_Atlantic_Northeast_Region = Middle_Atlantic_Northeast_Region, West_North_Central_Midwest_Region = West_North_Central_Midwest_Region, East_South_Central_South_Region = East_South_Central_South_Region, East_North_Central_Midwest_Region = East_North_Central_Midwest_Region, Pacific_West_Region = Pacific_West_Region, Puerto_Rico = Puerto_Rico, South_Atlantic_South_Region = South_Atlantic_South_Region, Mountain_West_Region = Mountain_West_Region, New_England_Northeast_Region = New_England_Northeast_Region)
if(all_values){
  return(div_list)
}else{
  return(div_list[[target_var]])
}
}
#Example of how to use this function
test_division <- subset_division(2022, target_var = 'Mountain_West_Region', all_values = FALSE)
test_division
```

*The project rubric mentions that all of these functions are supposed to be wrapped into one big function. I tried many times to make this function but I was not able to get it to work. Every one that I would write would return the wrong information or an error. The functions listed below are the closest I was able to get.*

*This is the first function for the years 2021 and 2022.*

```{r}
nested_function_21_22 <- function(year,...){
  return_date <- function(year = 2024){
  rds <- paste0(year, '_variables.rds')
  variables <- readRDS(rds)
  return(variables)
  }
  return_num_vars <- function(year, target_var = 'AGEP'){
  rds <- paste0(year, "_variables.rds")
  var_list <- readRDS(rds)
  PWGTP = c(var_list$variables$PWGTP$values$range$min, var_list$variables$PWGTP$values$range$max)
  AGEP = c(var_list$variables$AGEP$values$range$min, var_list$variables$AGEP$values$range$max)
  GASP = c(var_list$variables$GASP$values$range$min, var_list$variables$GASP$values$range$max)
  GRPIP = c(var_list$variables$GRPIP$values$range$min, var_list$variables$GRPIP$values$range$max)
  JWMNP = c(var_list$variables$JWMNP$values$range$min, var_list$variables$JWMNP$values$range$max)
  full_list <- list(AGEP = AGEP, GASP = GASP, GRPIP = GRPIP, JWMNP = JWMNP, PWGTP = PWGTP)
  num_list <- lapply(full_list, as.numeric, na.rm = TRUE)
  return(list(num_list['PWGTP'], num_list[target_var]))
  }
  return_JWAP <- function(year){
variables <- paste0(year, '_variables.rds')
variables_list <- readRDS(variables)
variables_df <- data.frame(
  data = unlist(variables_list$variables$JWAP$values$item),
  stringsAsFactors = FALSE
)

to_minutes <- function(time_str, am_pm){
  parts <- str_split(time_str, ':', simplify = TRUE)
  hours <- as.numeric(parts[, 1])
  minutes <- as.numeric(parts[, 2])
  
  is_pm <- str_detect(am_pm, 'p')
  
  hours <- case_when(
    is_pm & hours != 12 ~ hours + 12,
    !is_pm & hours == 12 ~ 0,
    TRUE                 ~ hours
  )
  return(hours * 60 + minutes)
}
processed <- variables_df |>
  mutate(
    start_time = str_extract(data, "^\\d{1,2}:\\d{2}"),
    start_ampm = str_extract(data, "^\\d{1,2}:\\d{2}\\s*[ap]\\.m\\."),
    end_time = str_extract(data, "\\d{1,2}:\\d{2}\\s*[ap]\\.m\\.$") %>% str_extract("^\\d{1,2}:\\d{2}"),
    end_ampm = str_extract(data, "\\d{1,2}:\\d{2}\\s*[ap]\\.m\\.$"),
    start_mins = to_minutes(start_time, start_ampm),
    end_mins = to_minutes(end_time, end_ampm),
    end_mins = if_else(end_mins < start_mins, end_mins + 1440, end_mins),
    midpoint_mins = (start_mins + end_mins) / 2,
    midpoint_mins = midpoint_mins %% 1440,
    midpoint_time = sprintf("%02d:%02d", floor(midpoint_mins / 60), round(midpoint_mins %% 60))
  ) |>
  select(midpoint_time)
print(processed)
  }
  return_JWDP <- function(year){
  variables <- paste0(year, '_variables.rds')
variables_list <- readRDS(variables)
variables_df <- data.frame(
  data = unlist(variables_list$variables$JWDP$values$item),
  stringsAsFactors = FALSE
)

to_minutes <- function(time_str, am_pm){
  parts <- str_split(time_str, ':', simplify = TRUE)
  hours <- as.numeric(parts[, 1])
  minutes <- as.numeric(parts[, 2])
  
  is_pm <- str_detect(am_pm, 'p')
  
  hours <- case_when(
    is_pm & hours != 12 ~ hours + 12,
    !is_pm & hours == 12 ~ 0,
    TRUE                 ~ hours
  )
  return(hours * 60 + minutes)
}
processed <- variables_df |>
  mutate(
    start_time = str_extract(data, "^\\d{1,2}:\\d{2}"),
    start_ampm = str_extract(data, "^\\d{1,2}:\\d{2}\\s*[ap]\\.m\\."),
    end_time = str_extract(data, "\\d{1,2}:\\d{2}\\s*[ap]\\.m\\.$") %>% str_extract("^\\d{1,2}:\\d{2}"),
    end_ampm = str_extract(data, "\\d{1,2}:\\d{2}\\s*[ap]\\.m\\.$"),
    start_mins = to_minutes(start_time, start_ampm),
    end_mins = to_minutes(end_time, end_ampm),
    end_mins = if_else(end_mins < start_mins, end_mins + 1440, end_mins),
    midpoint_mins = (start_mins + end_mins) / 2,
    midpoint_mins = midpoint_mins %% 1440,
    midpoint_time = sprintf("%02d:%02d", floor(midpoint_mins / 60), round(midpoint_mins %% 60))
  ) |>
  select(midpoint_time)
print(processed)
  }
  return_cat_vars <- function(year, target_var = 'SEX'){
rds <- paste0(year, '_variables.rds')
var_list <- readRDS(rds)
FER = factor(unlist(var_list$variables$FER))
HHL = factor(unlist(var_list$variables$HHL))
SCH = factor(unlist(var_list$variables$SCH))
SCHL = factor(unlist(var_list$variables$SCHL))
SEX = factor(unlist(var_list$variables$SEX))
var_list <- list(FER = FER, HHL = HHL, SCH = SCH, SCHL = SCHL, SEX = SEX)
return(var_list[[target_var]])
  }
  geography_level <- function(year, target_var = 'REGION', all_values = FALSE){
rds <- paste0(year, '_variables.rds')
vars_list <- readRDS(rds)
REGION = vars_list$variables$REGION
ST = vars_list$variables$ST
DIVISION = vars_list$variables$DIVISION
geo_list <- list(REGION = REGION, ST = ST, DIVISION = DIVISION)
if(all_values){
  return('ERROR')
}else{
  return(geo_list[[target_var]])
}
  }
  subset_region <- function(year, target_var = 'Northeast', all_values = FALSE){
rds <- paste0(year, '_variables.rds')
vars_list <- readRDS(rds)
Northeast = vars_list$variables$REGION$values$item$`1`
Puerto_Rico = vars_list$variables$REGION$values$item$`9`
South = vars_list$variables$REGION$values$item$`3`
West = vars_list$variables$REGION$values$item$`4`
Midwest = vars_list$variables$REGION$values$item$`2`
region_list <- list(Northeast = Northeast, Puerto_Rico = Puerto_Rico, South = South, West = West, Midwest = Midwest)
if(all_values){
  return(region_list)
}else{
  return(region_list[[target_var]])
}
  }
  subset_st <- function(year, target_var = 'North_Carolina', all_values = FALSE){
rds <- paste0(year, '_variables.rds')
vars_list <- readRDS(rds)
Tennessee = vars_list$variables$ST$values$item$`47`
Arizona = vars_list$variables$ST$values$item$`04`
Arkansas = vars_list$variables$ST$values$item$`05`
Utah = vars_list$variables$ST$values$item$`49`
Connecticut = vars_list$variables$ST$values$item$`09`
Vermont = vars_list$variables$ST$values$item$`50`
Mississippi = vars_list$variables$ST$values$item$`28`
North_Carolina = vars_list$variables$ST$values$item$`37`
Pennsylvania = vars_list$variables$ST$values$item$`42`
South_Carolina = vars_list$variables$ST$values$item$`45`
Virginia = vars_list$variables$ST$values$item$`51`
Hawaii = vars_list$variables$ST$values$item$`15`
Massachusetts = vars_list$variables$ST$values$item$`25`
Puerto_Rico = vars_list$variables$ST$values$item$`72`
Montana = vars_list$variables$ST$values$item$`30`
New_Hampshire = vars_list$variables$ST$values$item$`33`
New_Mexico = vars_list$variables$ST$values$item$`35`
Florida = vars_list$variables$ST$values$item$`12`
Texas = vars_list$variables$ST$values$item$`48`
Washington = vars_list$variables$ST$values$item$`53`
Georgia = vars_list$variables$ST$values$item$`13`
Indiana = vars_list$variables$ST$values$item$`18`
Kansas = vars_list$variables$ST$values$item$`20`
Michigan = vars_list$variables$ST$values$item$`26`
Nevada = vars_list$variables$ST$values$item$`32`
West_Virginia = vars_list$variables$ST$values$item$`54`
Wisconsin = vars_list$variables$ST$values$item$`55`
Minnesota = vars_list$variables$ST$values$item$`27`
Idaho = vars_list$variables$ST$values$item$`16`
Iowa = vars_list$variables$ST$values$item$`19`
Louisiana = vars_list$variables$ST$values$item$`22`
North_Dakota = vars_list$variables$ST$values$item$`38`
Alabama = vars_list$variables$ST$values$item$`01`
Alaska = vars_list$variables$ST$values$item$`02`
California = vars_list$variables$ST$values$item$`06`
Colorado = vars_list$variables$ST$values$item$`08`
District_of_Columbia = vars_list$variables$ST$values$item$`11`
Maine = vars_list$variables$ST$values$item$`23`
New_Jersey = vars_list$variables$ST$values$item$`34`
Ohio = vars_list$variables$ST$values$item$`39`
Delaware = vars_list$variables$ST$values$item$`10`
Wyoming = vars_list$variables$ST$values$item$`56`
Kentucky = vars_list$variables$ST$values$item$`21`
Maryland = vars_list$variables$ST$values$item$`24`
Rhode_Island = vars_list$variables$ST$values$item$`44`
South_Dakota = vars_list$variables$ST$values$item$`46`
Illinois = vars_list$variables$ST$values$item$`17`
Missouri = vars_list$variables$ST$values$item$`29`
Nebraska = vars_list$variables$ST$values$item$`31`
New_York = vars_list$variables$ST$values$item$`36`
Oklahoma = vars_list$variables$ST$values$item$`40`
Oregon = vars_list$variables$ST$values$item$`41`
st_list <- list(Tennessee = Tennessee, Arizona = Arizona, Arkansas = Arkansas, Utah = Utah, Connecticut = Connecticut, Vermont = Vermont, Mississippi = Mississippi, North_Carolina = North_Carolina, Pennsylvania = Pennsylvania, South_Carolina = South_Carolina, Virginia = Virginia, Hawaii = Hawaii, Massachusetts = Massachusetts, Puerto_Rico = Puerto_Rico, Montana = Montana, New_Hampshire = New_Hampshire, New_Mexico = New_Mexico, Florida = Florida, Texas = Texas, Washington = Washington, Georgia = Georgia, Indiana = Indiana, Kansas = Kansas, Michigan = Michigan, Nevada = Nevada, West_Virginia = West_Virginia, Wisconsin = Wisconsin, Minnesota = Minnesota, Idaho = Idaho, Iowa = Iowa, Louisiana = Louisiana, North_Dakota = North_Dakota, Alabama = Alabama, Alaska = Alaska, California = California, Colorado = Colorado, District_of_Columbia = District_of_Columbia, Maine = Maine, New_Jersey = New_Jersey, Ohio = Ohio, Delaware = Delaware, Wyoming = Wyoming, Kentucky = Kentucky, Maryland = Maryland, Rhode_Island = Rhode_Island, South_Dakota = South_Dakota, Illinois = Illinois, Missouri = Missouri, Nebraska = Nebraska, New_York = New_York, Oklahoma = Oklahoma, Oregon = Oregon)

if(all_values){
  return(st_list)
}else{
  return(st_list[[target_var]])
}
  }
  subset_division <- function(year, target_var = 'Puerto_Rico', all_values = FALSE){
rds <- paste0(year, '_variables.rds')
vars_list <- readRDS(rds)
West_South_Central_South_Region = vars_list$variables$DIVISION$values$item$`7`
Middle_Atlantic_Northeast_Region = vars_list$variables$DIVISION$values$item$`2`
West_North_Central_Midwest_Region = vars_list$variables$DIVISION$values$item$`4`
East_South_Central_South_Region = vars_list$variables$DIVISION$values$item$`6`
East_North_Central_Midwest_Region = vars_list$variables$DIVISION$values$item$`3`
Pacific_West_Region = vars_list$variables$DIVISION$values$item$`9`
Puerto_Rico = vars_list$variables$DIVISION$values$item$`0`
South_Atlantic_South_Region = vars_list$variables$DIVISION$values$item$`5`
Mountain_West_Region = vars_list$variables$DIVISION$values$item$`8`
New_England_Northeast_Region = vars_list$variables$DIVISION$values$item$`1`
div_list <- list(West_South_Central_South_Region = West_South_Central_South_Region, Middle_Atlantic_Northeast_Region = Middle_Atlantic_Northeast_Region, West_North_Central_Midwest_Region = West_North_Central_Midwest_Region, East_South_Central_South_Region = East_South_Central_South_Region, East_North_Central_Midwest_Region = East_North_Central_Midwest_Region, Pacific_West_Region = Pacific_West_Region, Puerto_Rico = Puerto_Rico, South_Atlantic_South_Region = South_Atlantic_South_Region, Mountain_West_Region = Mountain_West_Region, New_England_Northeast_Region = New_England_Northeast_Region)
if(all_values){
  return(div_list)
}else{
  return(div_list[[target_var]])
}
}

}

#This is how I was using this function
nest_test <- nested_function_21_22(2021, return_JWAP)
nest_test
```

*This is the second function for the years 2023 and 2024.*

```{r}
nested_function_23_24 <- function(year,...){
  return_date <- function(year = 2024){
  rds <- paste0(year, '_variables.rds')
  variables <- readRDS(rds)
  return(variables)
  }
  return_num_vars <- function(year, target_var = 'AGEP'){
  rds <- paste0(year, "_variables.rds")
  var_list <- readRDS(rds)
  PWGTP = c(var_list$variables$PWGTP$values$range$min, var_list$variables$PWGTP$values$range$max)
  AGEP = c(var_list$variables$AGEP$values$range$min, var_list$variables$AGEP$values$range$max)
  GASP = c(var_list$variables$GASP$values$range$min, var_list$variables$GASP$values$range$max)
  GRPIP = c(var_list$variables$GRPIP$values$range$min, var_list$variables$GRPIP$values$range$max)
  JWMNP = c(var_list$variables$JWMNP$values$range$min, var_list$variables$JWMNP$values$range$max)
  full_list <- list(AGEP = AGEP, GASP = GASP, GRPIP = GRPIP, JWMNP = JWMNP, PWGTP = PWGTP)
  num_list <- lapply(full_list, as.numeric, na.rm = TRUE)
  return(list(num_list['PWGTP'], num_list[target_var]))
  }
  return_JWAP <- function(year){
variables <- paste0(year, '_variables.rds')
variables_list <- readRDS(variables)
variables_df <- data.frame(
  data = unlist(variables_list$variables$JWAP$values$item),
  stringsAsFactors = FALSE
)

to_minutes <- function(time_str, am_pm){
  parts <- str_split(time_str, ':', simplify = TRUE)
  hours <- as.numeric(parts[, 1])
  minutes <- as.numeric(parts[, 2])
  
  is_pm <- str_detect(am_pm, 'p')
  
  hours <- case_when(
    is_pm & hours != 12 ~ hours + 12,
    !is_pm & hours == 12 ~ 0,
    TRUE                 ~ hours
  )
  return(hours * 60 + minutes)
}
processed <- variables_df |>
  mutate(
    start_time = str_extract(data, "^\\d{1,2}:\\d{2}"),
    start_ampm = str_extract(data, "^\\d{1,2}:\\d{2}\\s*[ap]\\.m\\."),
    end_time = str_extract(data, "\\d{1,2}:\\d{2}\\s*[ap]\\.m\\.$") %>% str_extract("^\\d{1,2}:\\d{2}"),
    end_ampm = str_extract(data, "\\d{1,2}:\\d{2}\\s*[ap]\\.m\\.$"),
    start_mins = to_minutes(start_time, start_ampm),
    end_mins = to_minutes(end_time, end_ampm),
    end_mins = if_else(end_mins < start_mins, end_mins + 1440, end_mins),
    midpoint_mins = (start_mins + end_mins) / 2,
    midpoint_mins = midpoint_mins %% 1440,
    midpoint_time = sprintf("%02d:%02d", floor(midpoint_mins / 60), round(midpoint_mins %% 60))
  ) |>
  select(midpoint_time)
print(processed)
  }
  return_JWDP <- function(year){
  variables <- paste0(year, '_variables.rds')
variables_list <- readRDS(variables)
variables_df <- data.frame(
  data = unlist(variables_list$variables$JWDP$values$item),
  stringsAsFactors = FALSE
)

to_minutes <- function(time_str, am_pm){
  parts <- str_split(time_str, ':', simplify = TRUE)
  hours <- as.numeric(parts[, 1])
  minutes <- as.numeric(parts[, 2])
  
  is_pm <- str_detect(am_pm, 'p')
  
  hours <- case_when(
    is_pm & hours != 12 ~ hours + 12,
    !is_pm & hours == 12 ~ 0,
    TRUE                 ~ hours
  )
  return(hours * 60 + minutes)
}
processed <- variables_df |>
  mutate(
    start_time = str_extract(data, "^\\d{1,2}:\\d{2}"),
    start_ampm = str_extract(data, "^\\d{1,2}:\\d{2}\\s*[ap]\\.m\\."),
    end_time = str_extract(data, "\\d{1,2}:\\d{2}\\s*[ap]\\.m\\.$") %>% str_extract("^\\d{1,2}:\\d{2}"),
    end_ampm = str_extract(data, "\\d{1,2}:\\d{2}\\s*[ap]\\.m\\.$"),
    start_mins = to_minutes(start_time, start_ampm),
    end_mins = to_minutes(end_time, end_ampm),
    end_mins = if_else(end_mins < start_mins, end_mins + 1440, end_mins),
    midpoint_mins = (start_mins + end_mins) / 2,
    midpoint_mins = midpoint_mins %% 1440,
    midpoint_time = sprintf("%02d:%02d", floor(midpoint_mins / 60), round(midpoint_mins %% 60))
  ) |>
  select(midpoint_time)
print(processed)
  }
  return_cat_vars <- function(year, target_var = 'SEX'){
rds <- paste0(year, '_variables.rds')
var_list <- readRDS(rds)
FER = factor(unlist(var_list$variables$FER))
HHL = factor(unlist(var_list$variables$HHL))
SCH = factor(unlist(var_list$variables$SCH))
SCHL = factor(unlist(var_list$variables$SCHL))
SEX = factor(unlist(var_list$variables$SEX))
var_list <- list(FER = FER, HHL = HHL, SCH = SCH, SCHL = SCHL, SEX = SEX)
return(var_list[[target_var]])
  }
  geography_level_2 <- function(year, target_var = 'REGION', all_values = FALSE){
rds <- paste0(year, '_variables.rds')
vars_list <- readRDS(rds)
REGION = vars_list$variables$REGION
STATE = vars_list$variables$STATE
DIVISION = vars_list$variables$DIVISION
geo_list <- list(REGION = REGION, STATE = STATE, DIVISION = DIVISION)
if(all_values){
  return('ERROR')
}else{
  return(geo_list[[target_var]])
}
  }
  subset_region <- function(year, target_var = 'Northeast', all_values = FALSE){
rds <- paste0(year, '_variables.rds')
vars_list <- readRDS(rds)
Northeast = vars_list$variables$REGION$values$item$`1`
Puerto_Rico = vars_list$variables$REGION$values$item$`9`
South = vars_list$variables$REGION$values$item$`3`
West = vars_list$variables$REGION$values$item$`4`
Midwest = vars_list$variables$REGION$values$item$`2`
region_list <- list(Northeast = Northeast, Puerto_Rico = Puerto_Rico, South = South, West = West, Midwest = Midwest)
if(all_values){
  return(region_list)
}else{
  return(region_list[[target_var]])
}
  }
  subset_state <- function(year, target_var = 'North_Carolina', all_values = FALSE){
rds <- paste0(year, '_variables.rds')
vars_list <- readRDS(rds)
Tennessee = vars_list$variables$STATE$values$item$`47`
Arizona = vars_list$variables$STATE$values$item$`04`
Arkansas = vars_list$variables$STATE$values$item$`05`
Utah = vars_list$variables$STATE$values$item$`49`
Connecticut = vars_list$variables$STATE$values$item$`09`
Vermont = vars_list$variables$STATE$values$item$`50`
Mississippi = vars_list$variables$STATE$values$item$`28`
North_Carolina = vars_list$variables$STATE$values$item$`37`
Pennsylvania = vars_list$variables$STATE$values$item$`42`
South_Carolina = vars_list$variables$STATE$values$item$`45`
Virginia = vars_list$variables$STATE$values$item$`51`
Hawaii = vars_list$variables$STATE$values$item$`15`
Massachusetts = vars_list$variables$STATE$values$item$`25`
Puerto_Rico = vars_list$variables$STATE$values$item$`72`
Montana = vars_list$variables$STATE$values$item$`30`
New_Hampshire = vars_list$variables$STATE$values$item$`33`
New_Mexico = vars_list$variables$STATE$values$item$`35`
Florida = vars_list$variables$STATE$values$item$`12`
Texas = vars_list$variables$STATE$values$item$`48`
Washington = vars_list$variables$STATE$values$item$`53`
Georgia = vars_list$variables$STATE$values$item$`13`
Indiana = vars_list$variables$STATE$values$item$`18`
Kansas = vars_list$variables$STATE$values$item$`20`
Michigan = vars_list$variables$STATE$values$item$`26`
Nevada = vars_list$variables$STATE$values$item$`32`
West_Virginia = vars_list$variables$STATE$values$item$`54`
Wisconsin = vars_list$variables$STATE$values$item$`55`
Minnesota = vars_list$variables$STATE$values$item$`27`
Idaho = vars_list$variables$STATE$values$item$`16`
Iowa = vars_list$variables$STATE$values$item$`19`
Louisiana = vars_list$variables$STATE$values$item$`22`
North_Dakota = vars_list$variables$STATE$values$item$`38`
Alabama = vars_list$variables$STATE$values$item$`01`
Alaska = vars_list$variables$STATE$values$item$`02`
California = vars_list$variables$STATE$values$item$`06`
Colorado = vars_list$variables$STATE$values$item$`08`
District_of_Columbia = vars_list$variables$STATE$values$item$`11`
Maine = vars_list$variables$STATE$values$item$`23`
New_Jersey = vars_list$variables$STATE$values$item$`34`
Ohio = vars_list$variables$STATE$values$item$`39`
Delaware = vars_list$variables$STATE$values$item$`10`
Wyoming = vars_list$variables$STATE$values$item$`56`
Kentucky = vars_list$variables$STATE$values$item$`21`
Maryland = vars_list$variables$STATE$values$item$`24`
Rhode_Island = vars_list$variables$STATE$values$item$`44`
South_Dakota = vars_list$variables$STATE$values$item$`46`
Illinois = vars_list$variables$STATE$values$item$`17`
Missouri = vars_list$variables$STATE$values$item$`29`
Nebraska = vars_list$variables$STATE$values$item$`31`
New_York = vars_list$variables$STATE$values$item$`36`
Oklahoma = vars_list$variables$STATE$values$item$`40`
Oregon = vars_list$variables$STATE$values$item$`41`
state_list <- list(Tennessee = Tennessee, Arizona = Arizona, Arkansas = Arkansas, Utah = Utah, Connecticut = Connecticut, Vermont = Vermont, Mississippi = Mississippi, North_Carolina = North_Carolina, Pennsylvania = Pennsylvania, South_Carolina = South_Carolina, Virginia = Virginia, Hawaii = Hawaii, Massachusetts = Massachusetts, Puerto_Rico = Puerto_Rico, Montana = Montana, New_Hampshire = New_Hampshire, New_Mexico = New_Mexico, Florida = Florida, Texas = Texas, Washington = Washington, Georgia = Georgia, Indiana = Indiana, Kansas = Kansas, Michigan = Michigan, Nevada = Nevada, West_Virginia = West_Virginia, Wisconsin = Wisconsin, Minnesota = Minnesota, Idaho = Idaho, Iowa = Iowa, Louisiana = Louisiana, North_Dakota = North_Dakota, Alabama = Alabama, Alaska = Alaska, California = California, Colorado = Colorado, District_of_Columbia = District_of_Columbia, Maine = Maine, New_Jersey = New_Jersey, Ohio = Ohio, Delaware = Delaware, Wyoming = Wyoming, Kentucky = Kentucky, Maryland = Maryland, Rhode_Island = Rhode_Island, South_Dakota = South_Dakota, Illinois = Illinois, Missouri = Missouri, Nebraska = Nebraska, New_York = New_York, Oklahoma = Oklahoma, Oregon = Oregon)
if(all_values){
  return(state_list)
}else{
  return(state_list[[target_var]])
}
  }
  subset_division <- function(year, target_var = 'Puerto_Rico', all_values = FALSE){
rds <- paste0(year, '_variables.rds')
vars_list <- readRDS(rds)
West_South_Central_South_Region = vars_list$variables$DIVISION$values$item$`7`
Middle_Atlantic_Northeast_Region = vars_list$variables$DIVISION$values$item$`2`
West_North_Central_Midwest_Region = vars_list$variables$DIVISION$values$item$`4`
East_South_Central_South_Region = vars_list$variables$DIVISION$values$item$`6`
East_North_Central_Midwest_Region = vars_list$variables$DIVISION$values$item$`3`
Pacific_West_Region = vars_list$variables$DIVISION$values$item$`9`
Puerto_Rico = vars_list$variables$DIVISION$values$item$`0`
South_Atlantic_South_Region = vars_list$variables$DIVISION$values$item$`5`
Mountain_West_Region = vars_list$variables$DIVISION$values$item$`8`
New_England_Northeast_Region = vars_list$variables$DIVISION$values$item$`1`
div_list <- list(West_South_Central_South_Region = West_South_Central_South_Region, Middle_Atlantic_Northeast_Region = Middle_Atlantic_Northeast_Region, West_North_Central_Midwest_Region = West_North_Central_Midwest_Region, East_South_Central_South_Region = East_South_Central_South_Region, East_North_Central_Midwest_Region = East_North_Central_Midwest_Region, Pacific_West_Region = Pacific_West_Region, Puerto_Rico = Puerto_Rico, South_Atlantic_South_Region = South_Atlantic_South_Region, Mountain_West_Region = Mountain_West_Region, New_England_Northeast_Region = New_England_Northeast_Region)
if(all_values){
  return(div_list)
}else{
  return(div_list[[target_var]])
}
  }
  
}

#This is how I was using this function.
nest_test_2 <- nested_function_23_24(2023, return_JWAP)
nest_test_2
```

## COMBINING YEARS

*Writing function to specify multiple years of survey data.*

*I am still not sure if this is correct but it's not returning an error, so at least it's functional. It returns a tibble that is made up of 4 lists of 525. I am not able to see any of the elements of these lists.*

```{r}
#Copying the function that returns a single year into the same code chunk as the new function so that it all can be run together smoothly
return_date <- function(year = 2024){
  rds <- paste0(year, '_variables.rds')
  variables <- readRDS(rds)
  return(variables)
}

#Function that returns multiple years
combine_years <- function(return_date, survey_years){
  results <- list()
   for(i in seq_along(survey_years)){
     results[[i]] <- return_date(survey_years[i])
   }
  tibble_results <- as_tibble(results, .name_repair = 'minimal')
  class(tibble_results) <- c("census", class(tibble_results))
  return(tibble_results)
}

#Testing it out
test_combine <- combine_years(return_date, c(2021, 2022, 2023, 2024))
test_combine
```

# PART TWO: WRITING A GENERIC FUNCTION FOR SUMMARIZING

## NUMERIC SUMMARIES

*This part didn't give me much trouble. I was getting a lot of errors about trying to do math with vectors of unequal length, but it turns out I had inputted one of the vectors wrong.*

```{r}
#Writing the function
summary.census <- function(year, numeric_var, cat_var){
  rds <- paste0(year, '_variables.rds')
  var_list <- readRDS(rds)
  
  PWGTP = c(var_list$variables$PWGTP$values$range$min, var_list$variables$PWGTP$values$range$max)
  AGEP = c(var_list$variables$AGEP$values$range$min, var_list$variables$AGEP$values$range$max)
  GASP = c(var_list$variables$GASP$values$range$min, var_list$variables$GASP$values$range$max)
  GRPIP = c(var_list$variables$GRPIP$values$range$min, var_list$variables$GRPIP$values$range$max)
  JWMNP = c(var_list$variables$JWMNP$values$range$min, var_list$variables$JWMNP$values$range$max)
  full_list <- list(PWGTP = PWGTP, AGEP = AGEP, GASP = GASP, GRPIP = GRPIP, JWMNP = JWMNP)
  num_list <- lapply(full_list, as.numeric, na.rm = TRUE)
  numeric_var <- num_list[[numeric_var]]
  weight_vector <- num_list[['PWGTP']]
  sample_mean <- sum(numeric_var*weight_vector)/sum(weight_vector)
  sample_std_dev <- sqrt(sum(numeric_var*numeric_var*weight_vector)/sum(weight_vector)-sample_mean*sample_mean)
  
  FER = factor(unlist(var_list$variables$FER$values$item))
  HHL = factor(unlist(var_list$variables$HHL$values$item))
  SCH = factor(unlist(var_list$variables$SCH$values$item))
  SCHL = factor(unlist(var_list$variables$SCHL$values$item))
  SEX = factor(unlist(var_list$variables$SEX$values$item))
  var_list <- list(FER = FER, HHL = HHL, SCH = SCH, SCHL = SCHL, SEX = SEX)
  cat_var <- var_list[cat_var]
  return(list(sample_mean, sample_std_dev, cat_var))
}

#A couple of examples of hwo to use these functions
summary_test <- summary.census(2021, numeric_var = 'JWMNP', cat_var = 'HHL')
summary_test

summary_test_2 <- summary.census(2023, numeric_var = 'AGEP', cat_var = 'SEX')
summary_test_2
```

### INVESTIGATION

*For this section, I wanted to investigate the differences in the means and standard deviations of the GASP and GRPIP variables, as well as the counts of the SCHL language for both. I will be using data from the year 2024.*

```{r}
FER_SCHL_summary <- summary.census(2024, numeric_var = 'GASP', cat_var = 'SCHL')
FER_SCHL_summary
GRPIP_SCHL_summary <- summary.census(2024, numeric_var = 'GRPIP', cat_var = 'SCHL')
GRPIP_SCHL_summary
```

*The means returned varied greatly The GASP mean was 9,998 and the GRPIP mean was around 99.99. I suspect that there is something altering the* *value of the GASP mean, as there is no reason for someone's monthly gas expenses to be in the 9,000s. When looking at the SCHL values, I found* *that the demographic with the most responses was the Doctorate one, with 24 people in this category. Generally, it seems like a great majority of* *the people responding to this survey had been through some amount of college.*

## GRAPHICAL SUMMARIES

*I was not able to figure out how to get this one to work.*

```{r}
combined_census <- combine_years(return_date, c(2021,2022,2023,2024))


boxplot.census <- function(year, categorical_var, numeric_var){
  categorical_var <- combined_census[[categorical_var]]
  numeric_var <- combined_census[[numeric_var]]

  boxplot_result <- ggplot(data = combined_census, aes(x = categorical_var, y = numeric_var, fill = categorical_var, weight = 'PWGTP')) +
    geom_boxplot()
  return(boxplot_result)
}

boxplot_test <- boxplot.census(2021, categorical_var = 'SEX', numeric_var = 'AGEP')
boxplot_test
```
